### 2019-01-17 Nomenclature/Nomenclature.tex

\nomenclature[A]{AM}{\textit{Amplitude Modulation}}
\nomenclature[A]{ARTM}{\textit{Advanced Range Telemetry Group}}
\nomenclature[A]{COFDM}{\textit{Coded Orthogonal Frequency Division Multiplexing}}
\nomenclature[A]{CPM}{\textit{Continuous Phase Modulation}}
\nomenclature[A]{CPFSK}{\textit{Continuous Phase Frequency Shift Keying}}
\nomenclature[A]{FM}{\textit{Frequency Modulation}}
\nomenclature[A]{PCM/FM}{\textit{Pulse Code Modulation/Frequency Modulation}}
\nomenclature[A]{MHz}{\textit{Megahertz}}
\nomenclature[A]{GHz}{\textit{Gigahertz}}
\nomenclature[A]{Hz}{\textit{Hertz}}
\nomenclature[A]{dB}{\textit{Decibel}}
\nomenclature[A]{NRZ}{\textit{Non-Return to Zero}}
\nomenclature[A]{SOQPSK}{\textit{Shaped Offset Quadrature Phase-Shift Keying}}
\nomenclature[A]{SOQPSK-TG}{\textit{Shaped Offset Quadrature Phase-Shift Keying Telemetry Group}}
\nomenclature[A]{SOQPSK-MIL}{\textit{Shaped Offset Quadrature Phase-Shift Keying Military}}
\nomenclature[A]{MIL-STD SOQPSK}{\textit{Military Standard Shaped Offset Quadrature Phase-Shift Keying }}
\nomenclature[A]{APSK}{\textit{Amplitude Phase-Shift Keying}}
\nomenclature[A]{PSK}{\textit{Phase-Shift Keying}}
\nomenclature[A]{M-PSK}{\textit{M-ary Phase-Shift Keying}}
\nomenclature[A]{BPSK}{\textit{Binary Phase-Shift Keying}}
\nomenclature[A]{OQPSK}{\textit{Offset Quadrature Phase-Shift Keying}}
\nomenclature[A]{PAM}{\textit{Pulse Amplitude Modulation}}
\nomenclature[A]{M-QAM}{\textit{M-ary Quadrature Amplitude Modulation}}
\nomenclature[A]{MIMO}{\textit{Multiple Input Multiple Output}}
\nomenclature[A]{SISO}{\textit{Single Input Single Output}}
\nomenclature[A]{MISO}{\textit{Multiple Input Single Output}}
\nomenclature[A]{RF}{\textit{Radio Frequency}}
\nomenclature[A]{IRIG}{\textit{Inter-Range Instrumentation Group}}
\nomenclature[A]{RCC}{\textit{Range Commanders Council}}
\nomenclature[A]{ISI}{\textit{Inter-Symbol Interference}}
\nomenclature[A]{PRD}{\textit{Perrins-Rice Decomposition}}
\nomenclature[A]{DBD}{\textit{DuoBinary Decomposition}}
\nomenclature[A]{XTCQM}{\textit{Cross-correlated Trellis Coded Modulation}}
\nomenclature[A]{I$\&$D}{\textit{Integrate and Dump}}
\nomenclature[A]{CM}{\textit{Constant Modulus}}
\nomenclature[A]{CMA}{\textit{Constant Modulus Algorithm}}
\nomenclature[A]{LOS}{\textit{Line Of Sight}}
\nomenclature[A]{NLOS}{\textit{Non-Line Of Sight}}
\nomenclature[A]{STC}{\textit{Space-Time Coding}}
\nomenclature[A]{STBC}{\textit{Space-Time Block Coding}}
\nomenclature[A]{PSP}{\textit{PerSurvivor Processing}}
\nomenclature[A]{GMSK}{\textit{Gaussian Minimum Shift Keying}}
\nomenclature[A]{MSK}{\textit{Minimum Shift Keying}}
\nomenclature[A]{LS}{\textit{Least Square}}
\nomenclature[A]{LMS}{\textit{Least Mean Square}}
\nomenclature[A]{RLS}{\textit{Recursive Least Square}}
\nomenclature[A]{GSM}{\textit{Global System for Mobile communication }}
\nomenclature[A]{BT}{\textit{Bandwidth Time product}}
\nomenclature[A]{GFSK}{\textit{Gaussian Frequency Shift Keying}}
\nomenclature[A]{SBPSK}{\textit{Shaped Binary Phase-Shift Keying}}
\nomenclature[A]{DVB-RCS2}{\textit{Digital Video Broadcasting - Return Channel via Satellite - Second Generation}}
\nomenclature[A]{FQPSK}{\textit{Feher-patented Quadrature Phase-Shift Keying}}
\nomenclature[A]{RC}{\textit{Raised Cosine}}
\nomenclature[A]{REC}{\textit{Rectangular}}
\nomenclature[A]{BCJR}{\textit{Bahl, Cocke, Jelinek and Raviv}}
\nomenclature[A]{MAP}{\textit{Maximum A Posteriori}}
\nomenclature[A]{MLSD}{\textit{Maximum Likelihood Sequence Detection}}
\nomenclature[A]{CA-MLSD}{\textit{Conventional Adaptive Maximum Likelihood Sequence Detection}}
\nomenclature[A]{NA-MLSD}{\textit{Non-Adaptive Maximum Likelihood Sequence Detection}}
\nomenclature[A]{MLSD}{\textit{Maximum Likelihood Sequence Detection}}
\nomenclature[A]{MLSE}{\textit{Maximum Likelihood Sequence Estimation}}
\nomenclature[A]{Mbps}{\textit{Megabits per second}}
\nomenclature[A]{MMSE}{\textit{Minimum Mean Square Error}}
\nomenclature[A]{MSE}{\textit{Mean Square Error}}
\nomenclature[A]{SMA}{\textit{Saint Martin Airport}}
\nomenclature[A]{TBA}{\textit{Toulouse-Blagnac Airport}}
\nomenclature[A]{V2V}{\textit{Vehicle To Vehicle}}
\nomenclature[A]{AWGN}{\textit{Additive White Gaussian Noise}}
\nomenclature[A]{SxS}{\textit{Sample by Sample}}
\nomenclature[A]{SNR}{\textit{Signal-to-Noise Ratio}}
\nomenclature[A]{SER}{\textit{Symbol Error Rate}}
\nomenclature[A]{BER}{\textit{Bit Error Rate}}
\nomenclature[A]{ASM}{\textit{Asynchronous Marker}}
\nomenclature[A]{INET}{\textit{Integrated Network Enhanced Telemetry}}
\nomenclature[A]{STBD}{\textit{Space-Time Block Decoder}}
\nomenclature[A]{ML}{\textit{Maximum Likelihood}}
\nomenclature[A]{ML-ASM}{\textit{Maximum Likelihood - Approximate Signal Model}}
\nomenclature[A]{OD}{\textit{Offset Decode}}
\nomenclature[A]{SOVA}{\textit{Soft Output Viterbi Algorithm}}
\nomenclature[A]{LLR}{\textit{Log-Likelihood Ratio}}
\nomenclature[A]{FIR}{\textit{Finite Impulse Response}}
\nomenclature[A]{Max-Log-MAP}{\textit{Maximum Logarithm Maximum A Posteriori}}
\nomenclature[A]{FEC}{\textit{Forward Error Correction}}
\nomenclature[A]{LDPC}{\textit{Low Parity Density Check}}
\nomenclature[A]{CFO}{\textit{Carrier Frequency Offset}}
\nomenclature[A]{US}{\textit{United States}}
\nomenclature[A]{DF}{\textit{Decision Feedback}}
\nomenclature[A]{I.I.D}{\textit{Independent and Identically Distributed}}
\nomenclature[A]{ML-ASM}{\textit{Maximum Likelihood - Approximate Signal Model}}
\nomenclature[A]{NASA's JPL}{\textit{National Aeronautics and Space Administration's Jet Propulsion Lab}}
%\nomenclature[N]{$x$}{Scalar}