### 2019-01-17 2-Chapters/0-Introduction/Introduction.tex
Okay, all the template is better for my need now. I have to write…

%!TEX root = ../thesis.tex

% First chapter begins here
%\adjustmtc
\chapter*{Introduction}
\addcontentsline{toc}{chapter}{Introduction}
\chaptermark{Introduction}
\minitoc
\label{chapter:Intro}
\graphicspath{{2-Chapters/1-Chapter/Images/}}

% Write miniTOC just after the title
%\startcontents[chapters]\vbox{\bf\Large Sommaire \hrule}
%\printcontents[chapters]{}{1}{}\vspace*{1pc}\hrule}

% ----------------------------------------------------------------------------
\section*{General context}

\indent
The work presented in this PhD thesis
Blabla

\section*{Contributions}

Blabla

\begin{itemize}
\item Blabla
\item Blabla
\item Blabla
\end{itemize}

\section*{Dissertation outline}
This dissertation is organized as follows.

Blabla