### 2019-01-17 3-Appendices/1-Appendix/appendix1.tex

%!TEX root = ../thesis.tex
% ******************************* Thesis Appendix A ****************************
\chapter{Appendix for Chapter \ref{chapter:1}}
\label{app:chapt1}

In this appendix, blabla.