### 2019-02-07 4-Listofpublications/Listofpublications.tex
Fix #4

%!TEX root = ../thesis.tex
% ******************************* Thesis Appendix A ****************************
\chapter*{List of Publications}
\addcontentsline{toc}{chapter}{List of Publications}

% \section*{International Journals}
% \renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
% \begin{etaremune}%[label={[RI-\arabic*]}]
% 	\item \textbf{L. Besson}, E. Kaufmann ``XXX'',, in \emph{XXXX}, 2018 (Major Revision)
% \end{etaremune}

\section*{International Conferences}
\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
\begin{etaremune}
\item
  \href{https://perso.crans.org/besson/articles/BBM__IEEE_WCNC_2019.pdf}{GNU
  Radio Implementation of MALIN: « Multi-Armed bandits Learning for
  Internet-of-things Networks »}, par \textbf{L. Besson},
  \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi},
  \href{https://moychris.wordpress.com/}{C. Moy}.
  \href{http://wcnc2019.ieee-wcnc.org/}{IEEE WCNC},
  Marrakech, Maroc, avril 2019.
\item
  \href{https://perso.crans.org/besson/articles/BBM__Demo_ICT_2018.pdf}{MALIN:
  « Multi-Arm bandit Learning for Iot Networks » with GRC: A TestBed
  Implementation and Demonstration that Learning Helps}, par \textbf{L.
  Besson}, \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi},
  \href{https://moychris.wordpress.com/}{C. Moy}. Démonstration
  présentée à ICT (\emph{International Conference on Communication}),
  Saint-Malo, France, juin 2018. Cf. \texttt{YouTu.be/HospLNQhcMk}
\item
  \href{https://hal.inria.fr/hal-01629733/document}{Multi-Player Bandits
  Revisited}, par \textbf{L. Besson} \&
  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
  Kaufmann}.
  \href{http://www.cs.cornell.edu/conferences/alt2018/accepted.html}{ALT}
  (\emph{Algorithmic Learning Theory}), Lanzarote, Espagne, avril 2018,
  \href{https://hal.inria.fr/hal-01629733}{HAL-01629733}.
\item
  \href{https://hal.inria.fr/hal-01705292/document}{Aggregation of
  Multi-Armed Bandits learning algorithms for Opportunistic Spectrum
  Access}, par \textbf{L. Besson},
  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
  Kaufmann} \& \href{https://moychris.wordpress.com/}{C. Moy}.
  \href{http://wcnc2018.ieee-wcnc.org/}{IEEE WCNC}, Barcelone, Espagne,
  avril 2018, \href{https://hal.inria.fr/hal-01705292}{HAL-01705292}.
\item
  \href{https://hal.inria.fr/hal-01575419/document}{Multi-Armed Bandit
  Learning in IoT Networks and non-stationary settings}, par
  \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, \textbf{L.
  Besson}, \href{https://moychris.wordpress.com/}{C. Moy},
  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
  Kaufmann} \& J. Palicot. \href{http://crowncom.org/2017/}{CrownCom}
  (\emph{Conference on Cognitive Radio Oriented Wireless Networks}),
  Lisbonne, Portugal, septembre 2017,
  \href{https://hal.inria.fr/hal-01575419}{HAL-01575419}, \textbf{prix
  du meilleur article}.
\end{etaremune}

% \section*{National Conferences}
% \renewcommand{\labelenumi}{[NC-\arabic{enumi}]}
% \begin{etaremune}
% 	\item \textbf{L. Besson}, E. Kaufmann, ``XXX'', in \emph{GRETSI}, Lille, France, Aug 2019
% \end{etaremune}

\section*{Other National Communications}
\renewcommand{\labelenumi}{[ONC-\arabic{enumi}]}
\begin{etaremune}
	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017
	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, Mar 2017
	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, Mar 2017
	\item	R. Bonnefoi, \textbf{L. Besson}, ``Multi-Armed Bandit Learning in IoT Networks'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016
\end{etaremune}

\section*{Other Publications}
\renewcommand{\labelenumi}{[OP-\arabic{enumi}]}
\begin{etaremune}
\item
	\href{https://perso.crans.org/besson/articles/BMBBM__IEEE_WCNC__2019.pdf}{Upper-Confidence
	Bound for Channel Selection in LPWA Networks with Retransmissions},
	par \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, \textbf{L. Besson}, J.
	Manco-Vasquez,
	\href{https://moychris.wordpress.com/}{C. Moy}. Septembre 2018.
\item
	\href{https://hal.inria.fr/hal-01840022/document}{SMPyBandits: an
	Open-Source Research Framework for Single and Multi-Players Multi-Arms
	Bandits (MAB) Algorithms in Python}, par \textbf{L. Besson},
	développement depuis octobre 2016 et encore actif,
	\href{https://hal.inria.fr/hal-01840022}{HAL-01840022}. Code sur
	\href{https://GitHub.com/SMPyBandits/SMPyBandits}{GitHub.com/SMPyBandits}
	(environ 40000 lignes), documentation sur
	\href{https://SMPyBandits.rtfd.io}{SMPyBandits.rtfd.io}.
\item
	\href{https://hal.inria.fr/hal-01736357/document}{What Doubling-Trick
	Can and Can't Do for Multi-Armed Bandits}, par \textbf{L. Besson} \&
	\href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
	Kaufmann}, septembre 2018,
	\href{https://hal.inria.fr/hal-01736357}{HAL-01736357}.
\item
	\href{https://hal.inria.fr/hal-01847480/document}{A Note on the Ei
	Function and a Useful Sum-Inequality}, par \textbf{L. Besson}, février
	2018 \href{https://hal.inria.fr/hal-01847480}{HAL-01847480}.
\end{etaremune}

% \section*{Patents}
% \renewcommand{\labelenumi}{[PT-\arabic{enumi}]}
% \begin{etaremune}
% \item \textbf{L. Besson}, E. Kaufmann. ``XXX'', (pending)
% \end{etaremune}