### 2019-02-12 3-Appendices/2-Appendix/appendix2.tex
Still fighting against my template...

%!TEX root = ../PhD_thesis__Lilian_Besson
% ******************************* Thesis Appendix B ****************************
\chapter{Publication: What Doubling-Trick Can and Can't Do for Multi-Armed Bandits}
\label{app:2:DoublingTricks}

See https://hal.inria.fr/hal-01736357

I cannot include it raw without more work on the maths, we have to write only the best versions of our results.

But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!

I need to try DoublingTrick + GLRklUCB, at least experimentally!
For the theoretical analysis, it will probably be a huge failure.