### 2019-02-12 Nomenclature/Nomenclature.tex

\nomenclature[A]{UCB}{\textit{Upper Confidence Bound} (object or algorithm)}
\nomenclature[A]{klUCB}{\textit{Kullback-Leibler Upper Confidence Bound} (algorithm)}
\nomenclature[A]{KL}{\textit{Kullback-Leibler} (divergence)}
\nomenclature[A]{MCTopM}{\textit{Musical-Chair on Top-M} (algorithm)}
\nomenclature[A]{RandTopM}{\textit{Random Hoping Musical-Chair on Top-M} (algorithm)}
\nomenclature[A]{GLR}{\textit{Generalized Likelihood Ratio}}
\nomenclature[A]{GLRT}{\textit{Generalized Likelihood Ratio Test}}