### 2019-02-13 Nomenclature/Nomenclature.tex
Baby front-page, it works

\nomenclature[A]{wrt}{\textit{with respect to}}
\nomenclature[A]{wlog}{\textit{with loss of generality}} % TODO not used?
\nomenclature[A]{iid}{\textit{identically and independently distributed}}

% Objects
\nomenclature[A]{USRP}{\textit{Universal Software Radio Peripheral}}

% Institution
\nomenclature[A]{ANR}{\textit{Agence Nationale de la Recherche} (French National Research Agency)}
\nomenclature[A]{ENS}{\textit{École Normale Supérieure}}
\nomenclature[A]{CNRS}{\textit{Centre National de la Recherche Scientifique}}
\nomenclature[A]{SCEE}{\textit{Signal Communication Électronique Embarquée} (research team in CentraleSupélec, campus of Rennes)}

% IoT networks terms
\nomenclature[A]{IoT}{\textit{Internet of Things}}
\nomenclature[A]{NB-IoT}{\textit{Narrow-Band Internet of Things}}
\nomenclature[A]{LPWAN}{\textit{Low-Power Wide-Area Network}}
\nomenclature[A]{CR}{\textit{Cognitive Radio}}
\nomenclature[A]{OSA}{\textit{Opportunistic Spectrum Access}}
\nomenclature[A]{SU}{\textit{Secondary User}}
\nomenclature[A]{PU}{\textit{Primary User}}
\nomenclature[A]{BST}{\textit{Base-Station} (or \textit{gateway})}
\nomenclature[A]{QPSK}{\textit{Quadrature Phase-Shift Keying}}

% Communication terms
\nomenclature[A]{PLR}{\textit{Packet Loss Ratio}}

% Algorithms or tools
\nomenclature[A]{UCB}{\textit{Upper Confidence Bound} (object or algorithm)}
\nomenclature[A]{klUCB}{\textit{Kullback-Leibler Upper Confidence Bound} (algorithm)}
\nomenclature[A]{TS}{\textit{Thompson Sampling} (algorithm)}
\nomenclature[A]{KL}{\textit{Kullback-Leibler} (divergence)}
\nomenclature[A]{MCTopM}{\textit{Musical-Chair on Top-M} (algorithm)}
\nomenclature[A]{RandTopM}{\textit{Random Hoping Musical-Chair on Top-M} (algorithm)}
\nomenclature[A]{GLR}{\textit{Generalized Likelihood Ratio}}
\nomenclature[A]{GLRT}{\textit{Generalized Likelihood Ratio Test}}