### 2019-03-25 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__onlyfig.gv.tex

\begin{tikzpicture}[>=latex,line join=bevel,]
%%
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (18391.0bp,11.5bp) node {Tree exploration for K=2 arms and M=2 players using Selfish KLUCB, for depth=3 : 68 leafs, 2 absorbing};
\end{scope}
  \node (105) at (11123.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (39) at (31639.0bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [0/1,0/1]]};
  \node (101) at (10387.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,1/2]]};
  \node (30) at (20228.0bp,128.0bp) [draw=red,ellipse] {[[1/1,0/1], [1/1,0/1]]};
  \node (206) at (29707.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,0/2]]};
  \node (13) at (3211.2bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [1/1,1/1]]};
  \node (181) at (25107.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (175) at (24003.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/2,0/1]]};
  \node (7) at (20978.0bp,215.0bp) [draw,ellipse] {[[0/1,0/0], [0/1,0/0]]};
  \node (73) at (5235.2bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,1/1]]};
  \node (241) at (36147.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,0/2]]};
  \node (159) at (21059.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [0/2,0/1]]};
  \node (93) at (8915.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,1/1]]};
  \node (184) at (25659.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (123) at (14435.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,2/2]]};
  \node (219) at (32099.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,0/1]]};
  \node (189) at (26579.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,2/2]]};
  \node (230) at (34123.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (198) at (28235.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/2,0/1]]};
  \node (22) at (11949.0bp,128.0bp) [draw=red,ellipse] {[[1/1,0/1], [1/1,0/1]]};
  \node (19) at (9285.2bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [1/1,1/1]]};
  \node (14) at (3579.2bp,128.0bp) [draw=red,ellipse] {[[0/1,1/1], [0/1,1/1]]};
  \node (8) at (26303.0bp,215.0bp) [draw,ellipse] {[[0/0,0/1], [0/1,0/0]]};
  \node (20) at (10664.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [1/1,1/1]]};
  \node (154) at (20139.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,0/2]]};
  \node (35) at (27039.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,1/1]]};
  \node (0) at (19381.0bp,302.0bp) [draw=green,ellipse] {[[0/0,0/0], [0/0,0/0]]};
  \node (160) at (21243.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,1/2]]};
  \node (171) at (23267.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (213) at (30995.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,0/1]]};
  \node (16) at (5787.2bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [1/1,1/1]]};
  \node (221) at (32467.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (236) at (35227.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,0/1]]};
  \node (109) at (11859.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,0/1]]};
  \node (107) at (11491.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (232) at (34491.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [0/2,0/1]]};
  \node (130) at (15723.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,1/2]]};
  \node (162) at (21611.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,0/2]]};
  \node (178) at (24555.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (194) at (27499.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,1/2]]};
  \node (204) at (29339.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,0/2]]};
  \node (62) at (3211.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,1/1]]};
  \node (15) at (4407.2bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [1/1,0/1]]};
  \node (103) at (10755.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,2/2]]};
  \node (143) at (18115.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (12) at (2382.2bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,1/1]]};
  \node (115) at (12963.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,2/2]]};
  \node (217) at (31731.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (145) at (18483.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,0/1]]};
  \node (216) at (31547.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (57) at (2291.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,2/2]]};
  \node (112) at (12411.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (27) at (17563.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,0/1]]};
  \node (242) at (36331.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,0/2]]};
  \node (152) at (19771.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (141) at (17747.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,0/2]]};
  \node (244) at (36699.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,0/2]]};
  \node (214) at (31179.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,0/1]]};
  \node (64) at (3579.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,1/1]]};
  \node (53) at (1555.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,1/1]]};
  \node (37) at (29523.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,0/1]]};
  \node (177) at (24371.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (98) at (9835.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,1/1]]};
  \node (131) at (15907.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,1/2]]};
  \node (50) at (1003.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,1/2]]};
  \node (191) at (26947.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,2/2]]};
  \node (84) at (7259.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,1/1]]};
  \node (163) at (21795.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,0/2]]};
  \node (166) at (22347.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (47) at (451.19bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,2/2]]};
  \node (91) at (8547.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,1/2]]};
  \node (170) at (23083.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,0/2]]};
  \node (60) at (2843.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,1/1]]};
  \node (183) at (25475.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/2,0/1]]};
  \node (89) at (8179.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,0/1]]};
  \node (106) at (11307.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,0/1]]};
  \node (146) at (18667.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,0/1]]};
  \node (111) at (12227.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,1/2]]};
  \node (52) at (1371.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,1/1]]};
  \node (210) at (30443.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (61) at (3027.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/1,1/2]]};
  \node (196) at (27867.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (174) at (23819.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (121) at (14067.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,1/1]]};
  \node (79) at (6339.2bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,2/2]]};
  \node (116) at (13147.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,1/2]]};
  \node (44) at (35779.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [0/1,0/1]]};
  \node (185) at (25843.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/2,0/1]]};
  \node (240) at (35963.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,0/1]]};
  \node (234) at (34859.0bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (151) at (19587.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (164) at (21979.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,1/2]]};
  \node (155) at (20323.0bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \node (87) at (7811.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (207) at (29891.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/2,0/1]]};
  \node (74) at (5419.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,2/2]]};
  \node (188) at (26395.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,2/2]]};
  \node (63) at (3395.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,1/1]]};
  \node (220) at (32283.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,0/1]]};
  \node (10) at (34309.0bp,215.0bp) [draw,ellipse] {[[0/0,1/1], [0/1,0/0]]};
  \node (129) at (15539.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,1/1]]};
  \node (38) at (30811.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [1/1,0/1]]};
  \node (137) at (17011.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (197) at (28051.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (33) at (25107.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,0/1]]};
  \node (132) at (16091.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,2/2]]};
  \node (128) at (15355.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,2/2]]};
  \node (205) at (29523.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (6) at (19769.0bp,215.0bp) [draw,ellipse] {[[1/1,0/0], [0/0,0/1]]};
  \node (228) at (33755.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,2/2]]};
  \node (153) at (19955.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/2,0/1]]};
  \node (85) at (7443.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,1/1]]};
  \node (158) at (20875.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/2,0/1]]};
  \node (193) at (27315.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (104) at (10939.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,1/2]]};
  \node (239) at (35779.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,1/2]]};
  \node (148) at (19035.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,0/1]]};
  \node (113) at (12595.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,2/2]]};
  \node (118) at (13515.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,2/2]]};
  \node (97) at (9651.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,1/1]]};
  \node (139) at (17379.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,0/2]]};
  \node (173) at (23635.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/2,0/1]]};
  \node (86) at (7627.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,0/1]]};
  \node (225) at (33203.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,1/2]]};
  \node (199) at (28419.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,0/2]]};
  \node (127) at (15171.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,1/2]]};
  \node (150) at (19403.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,1/2]]};
  \node (203) at (29155.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (218) at (31915.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [0/2,0/1]]};
  \node (25) at (14711.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [1/1,1/1]]};
  \node (229) at (33939.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/1,0/2]]};
  \node (133) at (16275.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,1/2]]};
  \node (36) at (27958.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [0/1,0/1]]};
  \node (209) at (30259.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/2,0/1]]};
  \node (5) at (18994.0bp,215.0bp) [draw,ellipse] {[[0/0,0/1], [0/0,0/1]]};
  \node (149) at (19219.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,0/1]]};
  \node (43) at (34765.0bp,128.0bp) [draw=red,ellipse] {[[0/1,1/1], [0/1,1/1]]};
  \node (125) at (14803.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,1/1]]};
  \node (223) at (32835.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,0/1]]};
  \node (110) at (12043.0bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \node (3) at (11491.0bp,215.0bp) [draw,ellipse] {[[0/0,0/1], [1/1,0/0]]};
  \node (67) at (4131.2bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,0/1]]};
  \node (78) at (6155.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,1/2]]};
  \node (165) at (22163.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,0/2]]};
  \node (94) at (9099.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,1/1]]};
  \node (117) at (13331.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/2,1/1]]};
  \node (237) at (35411.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/2,0/1]]};
  \node (122) at (14251.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,1/1]]};
  \node (51) at (1187.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,1/1]]};
  \node (49) at (819.19bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,1/1]]};
  \node (34) at (26303.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [0/1,1/1]]};
  \node (157) at (20691.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,0/1]]};
  \node (119) at (13699.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,1/2]]};
  \node (192) at (27131.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,2/2]]};
  \node (82) at (6891.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/1,1/2]]};
  \node (24) at (13424.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [1/1,1/1]]};
  \node (32) at (23083.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,0/1]]};
  \node (68) at (4315.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,0/1]]};
  \node (18) at (7811.2bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [1/1,0/1]]};
  \node (31) at (21243.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [0/1,0/1]]};
  \node (235) at (35043.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/2,0/1]]};
  \node (172) at (23451.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,0/2]]};
  \node (46) at (267.19bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,1/2]]};
  \node (208) at (30075.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (142) at (17931.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/2,0/1]]};
  \node (224) at (33019.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,1/2]]};
  \node (66) at (3947.2bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,0/1]]};
  \node (168) at (22715.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (226) at (33387.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,2/2]]};
  \node (227) at (33571.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,1/2]]};
  \node (202) at (28971.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (42) at (34309.0bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [0/1,0/1]]};
  \node (41) at (33478.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [0/1,1/1]]};
  \node (69) at (4499.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,0/1]]};
  \node (23) at (12595.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [0/1,1/1]]};
  \node (77) at (5971.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/1,1/2]]};
  \node (135) at (16643.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (238) at (35595.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [0/2,0/1]]};
  \node (161) at (21427.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,0/1]]};
  \node (76) at (5787.2bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,1/1]]};
  \node (138) at (17195.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (144) at (18299.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/2,0/1]]};
  \node (88) at (7995.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,0/1]]};
  \node (108) at (11675.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,0/1]]};
  \node (179) at (24739.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (45) at (83.193bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,2/2]]};
  \node (156) at (20507.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/2,0/1]]};
  \node (17) at (7075.2bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [1/1,1/1]]};
  \node (80) at (6523.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,1/1]]};
  \node (212) at (30811.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (140) at (17563.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (186) at (26027.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,1/2]]};
  \node (58) at (2475.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (55) at (1923.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (211) at (30627.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,0/1]]};
  \node (147) at (18851.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,0/1]]};
  \node (102) at (10571.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/2,1/1]]};
  \node (40) at (32467.0bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [1/1,0/1]]};
  \node (222) at (32651.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,0/1]]};
  \node (65) at (3763.2bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (92) at (8731.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,2/2]]};
  \node (124) at (14619.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,1/2]]};
  \node (180) at (24923.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,0/2]]};
  \node (72) at (5051.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,1/1]]};
  \node (1) at (3211.2bp,215.0bp) [draw,ellipse] {[[0/1,0/0], [0/0,1/1]]};
  \node (83) at (7075.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,1/1]]};
  \node (59) at (2659.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,1/2]]};
  \node (233) at (34675.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,0/1]]};
  \node (187) at (26211.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (29) at (19769.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [0/1,0/1]]};
  \node (200) at (28603.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (81) at (6707.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,1/1]]};
  \node (26) at (15999.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [0/1,1/1]]};
  \node (100) at (10203.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,2/2]]};
  \node (190) at (26763.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (201) at (28787.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (28) at (18943.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [1/1,0/1]]};
  \node (21) at (11491.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [1/1,0/1]]};
  \node (9) at (30811.0bp,215.0bp) [draw,ellipse] {[[0/1,0/0], [0/0,0/1]]};
  \node (176) at (24187.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (2) at (7075.2bp,215.0bp) [draw,ellipse] {[[0/0,1/1], [1/1,0/0]]};
  \node (95) at (9283.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,1/2]]};
  \node (75) at (5603.2bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,1/2]]};
  \node (11) at (1005.2bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [1/1,1/1]]};
  \node (169) at (22899.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (54) at (1739.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/1,1/2]]};
  \node (126) at (14987.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/1,1/2]]};
  \node (90) at (8363.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,2/2]]};
  \node (96) at (9467.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,1/1]]};
  \node (70) at (4683.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,0/1]]};
  \node (56) at (2107.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,2/2]]};
  \node (195) at (27683.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,1/2]]};
  \node (71) at (4867.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,1/1]]};
  \node (231) at (34307.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (99) at (10019.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/1,1/2]]};
  \node (120) at (13883.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,1/1]]};
  \node (4) at (14711.0bp,215.0bp) [draw,ellipse] {[[1/1,0/0], [0/0,1/1]]};
  \node (215) at (31363.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/1,0/2]]};
  \node (167) at (22531.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (243) at (36515.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,1/2]]};
  \node (182) at (25291.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,0/2]]};
  \node (114) at (12779.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,2/2]]};
  \node (48) at (635.19bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,1/1]]};
  \node (134) at (16459.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,2/2]]};
  \node (136) at (16827.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \draw [black,->] (39) ..controls (31692.0bp,107.22bp) and (31702.0bp,100.51bp)  .. (31710.0bp,92.0bp) .. controls (31716.0bp,85.469bp) and (31721.0bp,76.875bp)  .. (217);
  \draw (31790.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (36) ..controls (27838.0bp,118.26bp) and (27790.0bp,109.48bp)  .. (27749.0bp,92.0bp) .. controls (27734.0bp,85.571bp) and (27720.0bp,75.084bp)  .. (195);
  \draw (27810.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (25) ..controls (14579.0bp,114.93bp) and (14509.0bp,105.45bp)  .. (14483.0bp,92.0bp) .. controls (14472.0bp,86.08bp) and (14462.0bp,76.405bp)  .. (123);
  \draw (14554.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (37) ..controls (29459.0bp,97.451bp) and (29415.0bp,76.913bp)  .. (204);
  \draw (29458.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (18) ..controls (7916.8bp,117.38bp) and (7945.2bp,108.42bp)  .. (7967.2bp,92.0bp) .. controls (7975.2bp,86.062bp) and (7981.2bp,77.147bp)  .. (88);
  \draw (8047.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (33) ..controls (25273.0bp,122.71bp) and (25395.0bp,114.74bp)  .. (25436.0bp,92.0bp) .. controls (25446.0bp,86.408bp) and (25455.0bp,77.041bp)  .. (183);
  \draw (25469.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (16) ..controls (5885.1bp,114.44bp) and (5911.7bp,106.03bp)  .. (5933.2bp,92.0bp) .. controls (5942.5bp,85.923bp) and (5950.4bp,76.751bp)  .. (77);
  \draw (5965.7bp,84.5bp) node {$1/4$};
  \draw [black,->] (20) ..controls (10539.0bp,121.01bp) and (10487.0bp,112.13bp)  .. (10444.0bp,92.0bp) .. controls (10431.0bp,85.909bp) and (10419.0bp,75.953bp)  .. (101);
  \draw (10526.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (2) ..controls (7284.3bp,189.86bp) and (7584.5bp,155.18bp)  .. (18);
  \draw (7564.2bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (35) ..controls (27170.0bp,114.59bp) and (27239.0bp,105.07bp)  .. (27265.0bp,92.0bp) .. controls (27277.0bp,86.012bp) and (27288.0bp,76.199bp)  .. (193);
  \draw (27367.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (33) ..controls (25043.0bp,97.451bp) and (24999.0bp,76.913bp)  .. (180);
  \draw (25042.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (11) ..controls (746.55bp,125.93bp) and (343.19bp,120.68bp)  .. (209.19bp,92.0bp) .. controls (179.53bp,85.651bp) and (147.8bp,73.021bp)  .. (45);
  \draw (270.19bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (3) ..controls (11636.0bp,187.06bp) and (11789.0bp,158.81bp)  .. (22);
  \draw (11813.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (23) ..controls (12490.0bp,117.38bp) and (12461.0bp,108.42bp)  .. (12439.0bp,92.0bp) .. controls (12431.0bp,86.062bp) and (12425.0bp,77.147bp)  .. (112);
  \draw (12515.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (35) ..controls (26918.0bp,118.94bp) and (26869.0bp,110.13bp)  .. (26827.0bp,92.0bp) .. controls (26813.0bp,85.664bp) and (26799.0bp,75.32bp)  .. (190);
  \draw (26898.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (37) ..controls (29523.0bp,98.163bp) and (29523.0bp,82.548bp)  .. (205);
  \draw (29594.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (33) ..controls (25327.0bp,120.68bp) and (25601.0bp,110.43bp)  .. (25695.0bp,92.0bp) .. controls (25730.0bp,85.143bp) and (25769.0bp,72.053bp)  .. (185);
  \draw (25830.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (18281.0bp,293.4bp) and (8245.3bp,224.08bp)  .. (2);
  \draw (14025.0bp,258.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (40) ..controls (32392.0bp,108.95bp) and (32370.0bp,101.5bp)  .. (32351.0bp,92.0bp) .. controls (32337.0bp,84.707bp) and (32322.0bp,74.303bp)  .. (220);
  \draw (32407.0bp,84.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (10) ..controls (34081.0bp,190.71bp) and (33724.0bp,154.16bp)  .. (41);
  \draw (33999.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (6) ..controls (20101.0bp,194.85bp) and (20888.0bp,149.5bp)  .. (31);
  \draw (20660.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (19) ..controls (9478.8bp,120.96bp) and (9673.2bp,111.74bp)  .. (9741.2bp,92.0bp) .. controls (9762.8bp,85.72bp) and (9785.1bp,74.102bp)  .. (98);
  \draw (9847.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (44) ..controls (35590.0bp,124.38bp) and (35400.0bp,117.28bp)  .. (35240.0bp,92.0bp) .. controls (35192.0bp,84.334bp) and (35138.0bp,70.222bp)  .. (235);
  \draw (35256.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (33) ..controls (25266.0bp,120.21bp) and (25386.0bp,111.46bp)  .. (25488.0bp,92.0bp) .. controls (25529.0bp,84.212bp) and (25574.0bp,70.776bp)  .. (184);
  \draw (25620.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (15) ..controls (4288.5bp,117.67bp) and (4240.5bp,108.92bp)  .. (4200.2bp,92.0bp) .. controls (4184.8bp,85.536bp) and (4169.5bp,75.044bp)  .. (67);
  \draw (4261.2bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (29) ..controls (19877.0bp,112.58bp) and (19919.0bp,103.81bp)  .. (19933.0bp,92.0bp) .. controls (19940.0bp,86.147bp) and (19945.0bp,77.474bp)  .. (153);
  \draw (19960.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (36) ..controls (28099.0bp,118.46bp) and (28177.0bp,109.49bp)  .. (28204.0bp,92.0bp) .. controls (28213.0bp,86.28bp) and (28220.0bp,77.133bp)  .. (198);
  \draw (28234.0bp,84.5bp) node {$1/2$};
  \draw [red,->] (14) ..controls (3677.1bp,114.44bp) and (3703.7bp,106.03bp)  .. (3725.2bp,92.0bp) .. controls (3734.5bp,85.923bp) and (3742.4bp,76.751bp)  .. (65);
  \draw (3752.7bp,84.5bp) node {$1$};
  \draw [black,->] (10) ..controls (34641.0bp,194.83bp) and (35424.0bp,149.54bp)  .. (44);
  \draw (35198.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (33) ..controls (25107.0bp,98.163bp) and (25107.0bp,82.548bp)  .. (181);
  \draw (25178.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (24) ..controls (13245.0bp,120.78bp) and (13090.0bp,111.82bp)  .. (13035.0bp,92.0bp) .. controls (13019.0bp,85.991bp) and (13002.0bp,75.206bp)  .. (115);
  \draw (13106.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (40) ..controls (32573.0bp,117.38bp) and (32601.0bp,108.42bp)  .. (32623.0bp,92.0bp) .. controls (32631.0bp,86.062bp) and (32637.0bp,77.147bp)  .. (222);
  \draw (32702.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (4) ..controls (14406.0bp,193.83bp) and (13751.0bp,150.6bp)  .. (24);
  \draw (14210.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (13) ..controls (3211.2bp,98.163bp) and (3211.2bp,82.548bp)  .. (62);
  \draw (3272.2bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (25) ..controls (14496.0bp,124.1bp) and (14244.0bp,116.92bp)  .. (14158.0bp,92.0bp) .. controls (14137.0bp,85.931bp) and (14116.0bp,74.475bp)  .. (121);
  \draw (14229.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (44) ..controls (35613.0bp,122.94bp) and (35490.0bp,115.06bp)  .. (35449.0bp,92.0bp) .. controls (35439.0bp,86.401bp) and (35431.0bp,77.032bp)  .. (237);
  \draw (35531.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (2) ..controls (6769.6bp,193.83bp) and (6114.5bp,150.6bp)  .. (16);
  \draw (6565.2bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (9) ..controls (31038.0bp,190.67bp) and (31393.0bp,154.25bp)  .. (39);
  \draw (31340.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (31) ..controls (21054.0bp,124.38bp) and (20864.0bp,117.28bp)  .. (20704.0bp,92.0bp) .. controls (20656.0bp,84.334bp) and (20602.0bp,70.222bp)  .. (156);
  \draw (20720.0bp,84.5bp) node {$1/4$};
  \draw [red,->] (22) ..controls (11982.0bp,97.517bp) and (12001.0bp,79.93bp)  .. (110);
  \draw (12012.0bp,84.5bp) node {$1$};
  \draw [black,->] (12) ..controls (2327.7bp,107.54bp) and (2317.1bp,100.76bp)  .. (2309.2bp,92.0bp) .. controls (2303.3bp,85.478bp) and (2299.3bp,76.886bp)  .. (57);
  \draw (2380.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (41) ..controls (33344.0bp,115.41bp) and (33264.0bp,106.15bp)  .. (33194.0bp,92.0bp) .. controls (33153.0bp,83.582bp) and (33107.0bp,70.306bp)  .. (224);
  \draw (33210.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (21) ..controls (11633.0bp,123.74bp) and (11713.0bp,115.46bp)  .. (11779.0bp,92.0bp) .. controls (11797.0bp,85.626bp) and (11815.0bp,74.665bp)  .. (109);
  \draw (11825.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (41) ..controls (33619.0bp,118.46bp) and (33697.0bp,109.49bp)  .. (33724.0bp,92.0bp) .. controls (33733.0bp,86.28bp) and (33740.0bp,77.133bp)  .. (228);
  \draw (33799.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (39) ..controls (31819.0bp,121.39bp) and (31976.0bp,112.72bp)  .. (32031.0bp,92.0bp) .. controls (32047.0bp,86.062bp) and (32062.0bp,75.426bp)  .. (219);
  \draw (32118.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (37) ..controls (29689.0bp,122.71bp) and (29811.0bp,114.74bp)  .. (29852.0bp,92.0bp) .. controls (29862.0bp,86.408bp) and (29871.0bp,77.041bp)  .. (207);
  \draw (29885.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (8) ..controls (26660.0bp,195.69bp) and (27576.0bp,148.62bp)  .. (36);
  \draw (27295.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (27) ..controls (17783.0bp,120.68bp) and (18057.0bp,110.43bp)  .. (18151.0bp,92.0bp) .. controls (18186.0bp,85.143bp) and (18225.0bp,72.053bp)  .. (144);
  \draw (18286.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (19255.0bp,273.25bp) and (19135.0bp,246.99bp)  .. (5);
  \draw (19226.0bp,258.5bp) node {$1/4$};
  \draw [black,->] (13) ..controls (3362.5bp,116.15bp) and (3466.0bp,106.08bp)  .. (3504.2bp,92.0bp) .. controls (3521.3bp,85.675bp) and (3538.4bp,74.718bp)  .. (64);
  \draw (3604.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (36) ..controls (27904.0bp,107.54bp) and (27893.0bp,100.76bp)  .. (27885.0bp,92.0bp) .. controls (27879.0bp,85.478bp) and (27875.0bp,76.886bp)  .. (196);
  \draw (27956.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (38) ..controls (30953.0bp,123.74bp) and (31033.0bp,115.46bp)  .. (31099.0bp,92.0bp) .. controls (31117.0bp,85.626bp) and (31135.0bp,74.665bp)  .. (214);
  \draw (31145.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (2) ..controls (7075.2bp,185.16bp) and (7075.2bp,169.55bp)  .. (17);
  \draw (7141.2bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (25) ..controls (14915.0bp,119.43bp) and (15139.0bp,108.76bp)  .. (15218.0bp,92.0bp) .. controls (15250.0bp,85.155bp) and (15285.0bp,72.364bp)  .. (128);
  \draw (15329.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (35) ..controls (27226.0bp,125.61bp) and (27396.0bp,119.22bp)  .. (27452.0bp,92.0bp) .. controls (27464.0bp,86.384bp) and (27474.0bp,76.573bp)  .. (194);
  \draw (27487.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (9) ..controls (31168.0bp,195.69bp) and (32085.0bp,148.62bp)  .. (40);
  \draw (31794.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (38) ..controls (30645.0bp,122.94bp) and (30522.0bp,115.06bp)  .. (30481.0bp,92.0bp) .. controls (30471.0bp,86.401bp) and (30463.0bp,77.032bp)  .. (210);
  \draw (30563.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (31) ..controls (21045.0bp,123.06bp) and (20840.0bp,115.18bp)  .. (20770.0bp,92.0bp) .. controls (20752.0bp,86.013bp) and (20734.0bp,74.976bp)  .. (157);
  \draw (20831.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (0) ..controls (20439.0bp,293.13bp) and (29686.0bp,224.37bp)  .. (9);
  \draw (25862.0bp,258.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (36) ..controls (28009.0bp,106.92bp) and (28019.0bp,100.26bp)  .. (28027.0bp,92.0bp) .. controls (28034.0bp,85.47bp) and (28039.0bp,76.913bp)  .. (197);
  \draw (28119.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (11) ..controls (893.33bp,112.15bp) and (839.38bp,103.33bp)  .. (792.19bp,92.0bp) .. controls (756.05bp,83.324bp) and (716.15bp,70.574bp)  .. (48);
  \draw (807.69bp,84.5bp) node {$1/4$};
  \draw [black,->] (8) ..controls (26512.0bp,189.86bp) and (26813.0bp,155.18bp)  .. (35);
  \draw (26781.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (44) ..controls (36002.0bp,121.59bp) and (36283.0bp,112.09bp)  .. (36379.0bp,92.0bp) .. controls (36411.0bp,85.309bp) and (36446.0bp,72.45bp)  .. (243);
  \draw (36489.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (16) ..controls (5892.6bp,110.89bp) and (5941.9bp,102.17bp)  .. (5985.2bp,92.0bp) .. controls (6024.4bp,82.796bp) and (6068.0bp,69.925bp)  .. (78);
  \draw (6123.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (28) ..controls (18995.0bp,107.03bp) and (19005.0bp,100.36bp)  .. (19013.0bp,92.0bp) .. controls (19019.0bp,85.554bp) and (19024.0bp,77.115bp)  .. (148);
  \draw (19039.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (11) ..controls (838.77bp,123.64bp) and (706.24bp,115.96bp)  .. (594.19bp,92.0bp) .. controls (560.6bp,84.818bp) and (524.07bp,71.95bp)  .. (47);
  \draw (665.19bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (27) ..controls (17337.0bp,122.9bp) and (17047.0bp,114.56bp)  .. (16948.0bp,92.0bp) .. controls (16920.0bp,85.556bp) and (16890.0bp,73.136bp)  .. (136);
  \draw (17019.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (6) ..controls (19543.0bp,190.68bp) and (19189.0bp,154.3bp)  .. (28);
  \draw (19461.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (23) ..controls (12439.0bp,118.09bp) and (12329.0bp,108.54bp)  .. (12289.0bp,92.0bp) .. controls (12275.0bp,85.987bp) and (12261.0bp,75.584bp)  .. (111);
  \draw (12355.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (4) ..controls (14711.0bp,185.16bp) and (14711.0bp,169.55bp)  .. (25);
  \draw (14767.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (31) ..controls (21398.0bp,118.73bp) and (21512.0bp,109.61bp)  .. (21609.0bp,92.0bp) .. controls (21654.0bp,83.893bp) and (21704.0bp,70.26bp)  .. (163);
  \draw (21758.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (25) ..controls (14446.0bp,122.52bp) and (14019.0bp,112.89bp)  .. (13956.0bp,92.0bp) .. controls (13939.0bp,86.313bp) and (13923.0bp,75.449bp)  .. (120);
  \draw (14038.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (42) ..controls (34309.0bp,98.163bp) and (34308.0bp,82.548bp)  .. (231);
  \draw (34379.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (26) ..controls (15931.0bp,106.74bp) and (15905.0bp,99.058bp)  .. (15882.0bp,92.0bp) .. controls (15847.0bp,81.317bp) and (15808.0bp,69.051bp)  .. (130);
  \draw (15898.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (39) ..controls (31566.0bp,110.32bp) and (31555.0bp,102.83bp)  .. (31547.0bp,92.0bp) .. controls (31543.0bp,85.441bp) and (31541.0bp,77.158bp)  .. (216);
  \draw (31629.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (41) ..controls (33529.0bp,106.92bp) and (33539.0bp,100.26bp)  .. (33547.0bp,92.0bp) .. controls (33554.0bp,85.47bp) and (33559.0bp,76.913bp)  .. (227);
  \draw (33639.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (36) ..controls (28096.0bp,116.39bp) and (28180.0bp,107.19bp)  .. (28253.0bp,92.0bp) .. controls (28292.0bp,83.853bp) and (28336.0bp,70.708bp)  .. (199);
  \draw (28379.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (13) ..controls (3293.8bp,110.32bp) and (3317.3bp,102.66bp)  .. (3337.2bp,92.0bp) .. controls (3349.8bp,85.271bp) and (3362.0bp,75.397bp)  .. (63);
  \draw (3429.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (34) ..controls (26240.0bp,108.63bp) and (26229.0bp,101.61bp)  .. (26221.0bp,92.0bp) .. controls (26216.0bp,85.695bp) and (26213.0bp,77.483bp)  .. (187);
  \draw (26297.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (4) ..controls (14475.0bp,204.63bp) and (14127.0bp,190.76bp)  .. (13834.0bp,179.0bp) .. controls (13405.0bp,161.76bp) and (12891.0bp,140.98bp)  .. (23);
  \draw (13910.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (34) ..controls (26426.0bp,120.04bp) and (26477.0bp,111.18bp)  .. (26519.0bp,92.0bp) .. controls (26533.0bp,85.815bp) and (26546.0bp,75.718bp)  .. (189);
  \draw (26598.0bp,84.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (38) ..controls (30713.0bp,114.44bp) and (30687.0bp,106.03bp)  .. (30665.0bp,92.0bp) .. controls (30656.0bp,85.923bp) and (30648.0bp,76.751bp)  .. (211);
  \draw (30736.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (12) ..controls (2432.7bp,106.92bp) and (2443.1bp,100.26bp)  .. (2451.2bp,92.0bp) .. controls (2457.6bp,85.47bp) and (2462.5bp,76.913bp)  .. (58);
  \draw (2543.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (33) ..controls (25205.0bp,114.44bp) and (25232.0bp,106.03bp)  .. (25253.0bp,92.0bp) .. controls (25263.0bp,85.923bp) and (25270.0bp,76.751bp)  .. (182);
  \draw (25351.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (21) ..controls (11325.0bp,122.94bp) and (11202.0bp,115.06bp)  .. (11161.0bp,92.0bp) .. controls (11151.0bp,86.401bp) and (11143.0bp,77.032bp)  .. (105);
  \draw (11243.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (31) ..controls (21243.0bp,98.163bp) and (21243.0bp,82.548bp)  .. (160);
  \draw (21314.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (25) ..controls (14950.0bp,122.66bp) and (15282.0bp,113.85bp)  .. (15394.0bp,92.0bp) .. controls (15429.0bp,85.331bp) and (15466.0bp,72.321bp)  .. (129);
  \draw (15462.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (32) ..controls (23270.0bp,123.4bp) and (23455.0bp,115.78bp)  .. (23611.0bp,92.0bp) .. controls (23663.0bp,84.136bp) and (23721.0bp,69.81bp)  .. (174);
  \draw (23757.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (1) ..controls (3332.2bp,186.05bp) and (3443.8bp,160.27bp)  .. (14);
  \draw (3484.2bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (35) ..controls (27004.0bp,105.35bp) and (26995.0bp,98.856bp)  .. (26987.0bp,92.0bp) .. controls (26979.0bp,84.545bp) and (26971.0bp,75.378bp)  .. (191);
  \draw (27048.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (19) ..controls (9382.8bp,114.44bp) and (9409.0bp,106.02bp)  .. (9430.2bp,92.0bp) .. controls (9439.4bp,85.915bp) and (9447.1bp,76.741bp)  .. (96);
  \draw (9517.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (17) ..controls (7157.8bp,110.32bp) and (7181.3bp,102.66bp)  .. (7201.2bp,92.0bp) .. controls (7213.8bp,85.271bp) and (7226.0bp,75.397bp)  .. (84);
  \draw (7293.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (23) ..controls (12595.0bp,98.163bp) and (12595.0bp,82.548bp)  .. (113);
  \draw (12661.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (24) ..controls (13456.0bp,97.517bp) and (13475.0bp,79.93bp)  .. (118);
  \draw (13535.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (11) ..controls (775.46bp,124.3bp) and (476.36bp,117.25bp)  .. (375.19bp,92.0bp) .. controls (350.16bp,85.753bp) and (323.85bp,73.704bp)  .. (46);
  \draw (456.69bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (18) ..controls (7736.3bp,108.95bp) and (7714.2bp,101.5bp)  .. (7695.2bp,92.0bp) .. controls (7680.6bp,84.707bp) and (7665.8bp,74.303bp)  .. (86);
  \draw (7751.2bp,84.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (33) ..controls (24939.0bp,125.15bp) and (24802.0bp,118.01bp)  .. (24687.0bp,92.0bp) .. controls (24656.0bp,85.044bp) and (24623.0bp,72.462bp)  .. (178);
  \draw (24748.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (13) ..controls (3075.8bp,119.8bp) and (3002.9bp,111.12bp)  .. (2941.2bp,92.0bp) .. controls (2919.0bp,85.136bp) and (2895.8bp,73.527bp)  .. (60);
  \draw (3022.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (6) ..controls (19915.0bp,187.06bp) and (20067.0bp,158.81bp)  .. (30);
  \draw (20091.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (16) ..controls (5593.1bp,120.89bp) and (5398.3bp,111.63bp)  .. (5330.2bp,92.0bp) .. controls (5308.4bp,85.718bp) and (5285.8bp,74.099bp)  .. (73);
  \draw (5391.2bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (24) ..controls (13392.0bp,97.517bp) and (13373.0bp,79.93bp)  .. (117);
  \draw (13399.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (7) ..controls (21390.0bp,197.35bp) and (22643.0bp,146.79bp)  .. (32);
  \draw (22167.0bp,171.5bp) node {$1$};
  \draw [black,->] (32) ..controls (22985.0bp,114.44bp) and (22959.0bp,106.03bp)  .. (22937.0bp,92.0bp) .. controls (22928.0bp,85.923bp) and (22920.0bp,76.751bp)  .. (169);
  \draw (23008.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (11) ..controls (894.24bp,113.39bp) and (851.7bp,104.6bp)  .. (837.19bp,92.0bp) .. controls (830.46bp,86.15bp) and (826.24bp,77.586bp)  .. (49);
  \draw (918.69bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (37) ..controls (29743.0bp,120.68bp) and (30017.0bp,110.43bp)  .. (30111.0bp,92.0bp) .. controls (30146.0bp,85.143bp) and (30185.0bp,72.053bp)  .. (209);
  \draw (30246.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (10) ..controls (34454.0bp,186.97bp) and (34605.0bp,158.83bp)  .. (43);
  \draw (34630.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (25) ..controls (14658.0bp,107.25bp) and (14647.0bp,100.55bp)  .. (14639.0bp,92.0bp) .. controls (14633.0bp,85.48bp) and (14629.0bp,76.888bp)  .. (124);
  \draw (14710.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (21) ..controls (11393.0bp,114.44bp) and (11367.0bp,106.03bp)  .. (11345.0bp,92.0bp) .. controls (11336.0bp,85.923bp) and (11328.0bp,76.751bp)  .. (106);
  \draw (11416.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (19) ..controls (9055.1bp,124.53bp) and (8754.7bp,117.7bp)  .. (8653.2bp,92.0bp) .. controls (8628.7bp,85.79bp) and (8603.0bp,73.821bp)  .. (91);
  \draw (8734.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (17) ..controls (7075.2bp,98.163bp) and (7075.2bp,82.548bp)  .. (83);
  \draw (7136.2bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (17) ..controls (7011.2bp,97.451bp) and (6966.8bp,76.913bp)  .. (82);
  \draw (7010.7bp,84.5bp) node {$1/2$};
  \draw [black,->] (20) ..controls (10485.0bp,120.78bp) and (10330.0bp,111.82bp)  .. (10275.0bp,92.0bp) .. controls (10259.0bp,85.991bp) and (10242.0bp,75.206bp)  .. (100);
  \draw (10346.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (24) ..controls (13299.0bp,121.01bp) and (13247.0bp,112.13bp)  .. (13204.0bp,92.0bp) .. controls (13191.0bp,85.909bp) and (13179.0bp,75.953bp)  .. (116);
  \draw (13286.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (31) ..controls (21466.0bp,121.59bp) and (21747.0bp,112.09bp)  .. (21843.0bp,92.0bp) .. controls (21875.0bp,85.309bp) and (21910.0bp,72.45bp)  .. (164);
  \draw (21953.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (18) ..controls (7811.2bp,98.163bp) and (7811.2bp,82.548bp)  .. (87);
  \draw (7887.2bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (23) ..controls (12686.0bp,112.09bp) and (12710.0bp,104.12bp)  .. (12731.0bp,92.0bp) .. controls (12742.0bp,85.614bp) and (12752.0bp,76.034bp)  .. (114);
  \draw (12805.0bp,84.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (1) ..controls (3211.2bp,185.16bp) and (3211.2bp,169.55bp)  .. (13);
  \draw (3267.2bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (44) ..controls (35937.0bp,118.39bp) and (36048.0bp,108.95bp)  .. (36087.0bp,92.0bp) .. controls (36101.0bp,86.054bp) and (36114.0bp,75.795bp)  .. (241);
  \draw (36126.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (42) ..controls (34466.0bp,118.48bp) and (34577.0bp,109.06bp)  .. (34616.0bp,92.0bp) .. controls (34630.0bp,86.053bp) and (34643.0bp,75.794bp)  .. (233);
  \draw (34700.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (2) ..controls (6601.3bp,198.9bp) and (4912.6bp,145.1bp)  .. (15);
  \draw (5967.2bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (32) ..controls (22941.0bp,124.0bp) and (22861.0bp,115.76bp)  .. (22794.0bp,92.0bp) .. controls (22776.0bp,85.637bp) and (22758.0bp,74.676bp)  .. (168);
  \draw (22855.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (0) ..controls (18717.0bp,297.82bp) and (15204.0bp,280.94bp)  .. (12345.0bp,266.0bp) .. controls (8608.2bp,246.48bp) and (4011.3bp,220.53bp)  .. (1);
  \draw (12416.0bp,258.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (19730.0bp,282.43bp) and (20604.0bp,235.94bp)  .. (7);
  \draw (20291.0bp,258.5bp) node {$1/4$};
  \draw [black,->] (19) ..controls (9173.5bp,113.58bp) and (9130.7bp,104.8bp)  .. (9116.2bp,92.0bp) .. controls (9109.5bp,86.13bp) and (9105.5bp,77.562bp)  .. (94);
  \draw (9197.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (27) ..controls (17395.0bp,125.15bp) and (17258.0bp,118.01bp)  .. (17143.0bp,92.0bp) .. controls (17112.0bp,85.044bp) and (17079.0bp,72.462bp)  .. (137);
  \draw (17204.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [red,->] (43) ..controls (34798.0bp,97.517bp) and (34817.0bp,79.93bp)  .. (234);
  \draw (34828.0bp,84.5bp) node {$1$};
  \draw [black,->] (11) ..controls (1103.6bp,114.7bp) and (1130.0bp,106.23bp)  .. (1151.2bp,92.0bp) .. controls (1160.3bp,85.921bp) and (1167.8bp,76.748bp)  .. (51);
  \draw (1238.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (41) ..controls (33424.0bp,107.54bp) and (33413.0bp,100.76bp)  .. (33405.0bp,92.0bp) .. controls (33399.0bp,85.478bp) and (33395.0bp,76.886bp)  .. (226);
  \draw (33476.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (32) ..controls (23210.0bp,116.39bp) and (23276.0bp,107.53bp)  .. (23332.0bp,92.0bp) .. controls (23359.0bp,84.608bp) and (23388.0bp,72.613bp)  .. (172);
  \draw (23457.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (38) ..controls (30811.0bp,98.163bp) and (30811.0bp,82.548bp)  .. (212);
  \draw (30882.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (12) ..controls (2206.5bp,119.52bp) and (2056.5bp,109.98bp)  .. (2003.2bp,92.0bp) .. controls (1984.8bp,85.786bp) and (1966.2bp,74.584bp)  .. (55);
  \draw (2074.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (9) ..controls (30811.0bp,185.16bp) and (30811.0bp,169.55bp)  .. (38);
  \draw (30877.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (27) ..controls (17563.0bp,98.163bp) and (17563.0bp,82.548bp)  .. (140);
  \draw (17634.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (18514.0bp,291.66bp) and (12413.0bp,225.93bp)  .. (3);
  \draw (15979.0bp,258.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (10) ..controls (34309.0bp,185.16bp) and (34309.0bp,169.55bp)  .. (42);
  \draw (34385.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (44) ..controls (35934.0bp,118.73bp) and (36048.0bp,109.61bp)  .. (36145.0bp,92.0bp) .. controls (36190.0bp,83.893bp) and (36240.0bp,70.26bp)  .. (242);
  \draw (36294.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (26) ..controls (16186.0bp,125.61bp) and (16356.0bp,119.22bp)  .. (16412.0bp,92.0bp) .. controls (16424.0bp,86.384bp) and (16434.0bp,76.573bp)  .. (134);
  \draw (16492.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (16) ..controls (6009.2bp,121.27bp) and (6287.4bp,111.5bp)  .. (6383.2bp,92.0bp) .. controls (6416.2bp,85.282bp) and (6451.9bp,72.424bp)  .. (80);
  \draw (6450.7bp,84.5bp) node {$1/4$};
  \draw [black,->] (16) ..controls (5506.1bp,123.2bp) and (5017.1bp,114.16bp)  .. (4946.2bp,92.0bp) .. controls (4927.8bp,86.266bp) and (4909.6bp,75.137bp)  .. (71);
  \draw (5027.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (15) ..controls (4231.6bp,119.42bp) and (4081.5bp,109.83bp)  .. (4028.2bp,92.0bp) .. controls (4009.6bp,85.782bp) and (3990.8bp,74.58bp)  .. (66);
  \draw (4099.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (34) ..controls (26172.0bp,114.59bp) and (26103.0bp,105.07bp)  .. (26077.0bp,92.0bp) .. controls (26065.0bp,86.012bp) and (26055.0bp,76.199bp)  .. (186);
  \draw (26143.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (1) ..controls (2983.7bp,190.67bp) and (2628.5bp,154.25bp)  .. (12);
  \draw (2923.2bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (12) ..controls (2262.8bp,118.04bp) and (2214.6bp,109.27bp)  .. (2174.2bp,92.0bp) .. controls (2159.1bp,85.558bp) and (2144.3bp,75.069bp)  .. (56);
  \draw (2235.2bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (40) ..controls (32467.0bp,98.163bp) and (32467.0bp,82.548bp)  .. (221);
  \draw (32543.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (32) ..controls (23278.0bp,121.15bp) and (23474.0bp,112.03bp)  .. (23542.0bp,92.0bp) .. controls (23564.0bp,85.73bp) and (23586.0bp,74.112bp)  .. (173);
  \draw (23592.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (31) ..controls (21401.0bp,118.39bp) and (21512.0bp,108.95bp)  .. (21551.0bp,92.0bp) .. controls (21565.0bp,86.054bp) and (21578.0bp,75.795bp)  .. (162);
  \draw (21590.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (33) ..controls (24980.0bp,116.52bp) and (24914.0bp,107.66bp)  .. (24857.0bp,92.0bp) .. controls (24831.0bp,84.628bp) and (24802.0bp,72.633bp)  .. (179);
  \draw (24928.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (37) ..controls (29682.0bp,120.21bp) and (29802.0bp,111.46bp)  .. (29904.0bp,92.0bp) .. controls (29945.0bp,84.212bp) and (29990.0bp,70.776bp)  .. (208);
  \draw (30036.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (27) ..controls (17499.0bp,97.451bp) and (17455.0bp,76.913bp)  .. (139);
  \draw (17498.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (27) ..controls (17729.0bp,122.71bp) and (17851.0bp,114.74bp)  .. (17892.0bp,92.0bp) .. controls (17902.0bp,86.408bp) and (17911.0bp,77.041bp)  .. (142);
  \draw (17925.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (11) ..controls (1004.5bp,98.163bp) and (1004.1bp,82.548bp)  .. (50);
  \draw (1076.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (31) ..controls (21077.0bp,122.94bp) and (20954.0bp,115.06bp)  .. (20913.0bp,92.0bp) .. controls (20903.0bp,86.401bp) and (20895.0bp,77.032bp)  .. (158);
  \draw (20995.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (26) ..controls (15946.0bp,107.25bp) and (15935.0bp,100.55bp)  .. (15927.0bp,92.0bp) .. controls (15921.0bp,85.48bp) and (15917.0bp,76.888bp)  .. (131);
  \draw (15998.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (29) ..controls (19627.0bp,124.08bp) and (19547.0bp,115.84bp)  .. (19481.0bp,92.0bp) .. controls (19464.0bp,85.639bp) and (19446.0bp,74.679bp)  .. (150);
  \draw (19542.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (16) ..controls (5956.2bp,125.46bp) and (6093.6bp,118.44bp)  .. (6209.2bp,92.0bp) .. controls (6239.5bp,85.065bp) and (6272.2bp,72.483bp)  .. (79);
  \draw (6318.2bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (19) ..controls (9284.5bp,98.163bp) and (9284.1bp,82.548bp)  .. (95);
  \draw (9355.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (28) ..controls (18890.0bp,107.25bp) and (18879.0bp,100.55bp)  .. (18871.0bp,92.0bp) .. controls (18865.0bp,85.48bp) and (18861.0bp,76.888bp)  .. (147);
  \draw (18942.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (37) ..controls (29396.0bp,116.52bp) and (29330.0bp,107.66bp)  .. (29273.0bp,92.0bp) .. controls (29247.0bp,84.628bp) and (29218.0bp,72.633bp)  .. (203);
  \draw (29344.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (39) ..controls (31562.0bp,107.84bp) and (31532.0bp,99.997bp)  .. (31506.0bp,92.0bp) .. controls (31475.0bp,82.408bp) and (31440.0bp,70.298bp)  .. (215);
  \draw (31522.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (37) ..controls (29266.0bp,125.14bp) and (28870.0bp,118.94bp)  .. (28737.0bp,92.0bp) .. controls (28706.0bp,85.566bp) and (28672.0bp,72.787bp)  .. (200);
  \draw (28798.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (12) ..controls (2523.8bp,118.67bp) and (2602.4bp,109.74bp)  .. (2629.2bp,92.0bp) .. controls (2637.8bp,86.27bp) and (2644.4bp,77.121bp)  .. (59);
  \draw (2659.7bp,84.5bp) node {$1/2$};
  \draw [black,->] (38) ..controls (30909.0bp,114.44bp) and (30936.0bp,106.03bp)  .. (30957.0bp,92.0bp) .. controls (30967.0bp,85.923bp) and (30974.0bp,76.751bp)  .. (213);
  \draw (31034.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (27) ..controls (17661.0bp,114.44bp) and (17688.0bp,106.03bp)  .. (17709.0bp,92.0bp) .. controls (17719.0bp,85.923bp) and (17726.0bp,76.751bp)  .. (141);
  \draw (17807.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (20189.0bp,291.08bp) and (25445.0bp,226.54bp)  .. (8);
  \draw (23338.0bp,258.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (42) ..controls (34407.0bp,114.44bp) and (34433.0bp,106.02bp)  .. (34454.0bp,92.0bp) .. controls (34463.0bp,85.915bp) and (34471.0bp,76.741bp)  .. (232);
  \draw (34541.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (3) ..controls (11264.0bp,190.68bp) and (10910.0bp,154.3bp)  .. (20);
  \draw (11183.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (4) ..controls (15017.0bp,193.83bp) and (15672.0bp,150.6bp)  .. (26);
  \draw (15498.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (19) ..controls (9118.3bp,123.92bp) and (8984.9bp,116.34bp)  .. (8872.2bp,92.0bp) .. controls (8839.0bp,84.841bp) and (8803.0bp,71.973bp)  .. (92);
  \draw (8943.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (34) ..controls (26355.0bp,107.03bp) and (26365.0bp,100.36bp)  .. (26373.0bp,92.0bp) .. controls (26379.0bp,85.554bp) and (26384.0bp,77.115bp)  .. (188);
  \draw (26449.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (15) ..controls (4353.6bp,107.25bp) and (4343.1bp,100.55bp)  .. (4335.2bp,92.0bp) .. controls (4329.1bp,85.48bp) and (4324.8bp,76.888bp)  .. (68);
  \draw (4406.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (15) ..controls (4475.0bp,106.68bp) and (4500.4bp,99.009bp)  .. (4523.2bp,92.0bp) .. controls (4558.2bp,81.245bp) and (4597.5bp,68.978bp)  .. (70);
  \draw (4653.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (16) ..controls (5629.5bp,118.39bp) and (5518.8bp,108.95bp)  .. (5479.2bp,92.0bp) .. controls (5465.3bp,86.054bp) and (5452.1bp,75.795bp)  .. (74);
  \draw (5550.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (17) ..controls (7226.5bp,116.15bp) and (7330.0bp,106.08bp)  .. (7368.2bp,92.0bp) .. controls (7385.3bp,85.675bp) and (7402.4bp,74.718bp)  .. (85);
  \draw (7468.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (1) ..controls (2787.2bp,197.66bp) and (1457.8bp,146.44bp)  .. (11);
  \draw (2307.2bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (20) ..controls (10766.0bp,112.21bp) and (10806.0bp,103.88bp)  .. (10840.0bp,92.0bp) .. controls (10862.0bp,84.484bp) and (10885.0bp,73.057bp)  .. (104);
  \draw (10944.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (44) ..controls (35681.0bp,114.44bp) and (35655.0bp,106.03bp)  .. (35633.0bp,92.0bp) .. controls (35624.0bp,85.923bp) and (35616.0bp,76.751bp)  .. (238);
  \draw (35704.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (33) ..controls (24850.0bp,125.14bp) and (24454.0bp,118.94bp)  .. (24321.0bp,92.0bp) .. controls (24290.0bp,85.566bp) and (24256.0bp,72.787bp)  .. (176);
  \draw (24382.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (28) ..controls (19010.0bp,106.63bp) and (19036.0bp,98.961bp)  .. (19058.0bp,92.0bp) .. controls (19093.0bp,81.173bp) and (19133.0bp,68.905bp)  .. (149);
  \draw (19184.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (25) ..controls (14537.0bp,118.77bp) and (14390.0bp,108.88bp)  .. (14337.0bp,92.0bp) .. controls (14317.0bp,85.688bp) and (14297.0bp,74.332bp)  .. (122);
  \draw (14398.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (11) ..controls (1237.0bp,125.83bp) and (1542.0bp,120.24bp)  .. (1644.2bp,92.0bp) .. controls (1666.1bp,85.955bp) and (1688.6bp,74.348bp)  .. (54);
  \draw (1695.7bp,84.5bp) node {$1/4$};
  \draw [black,->] (25) ..controls (14763.0bp,107.03bp) and (14773.0bp,100.36bp)  .. (14781.0bp,92.0bp) .. controls (14787.0bp,85.554bp) and (14792.0bp,77.115bp)  .. (125);
  \draw (14862.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (42) ..controls (34197.0bp,113.58bp) and (34155.0bp,104.8bp)  .. (34140.0bp,92.0bp) .. controls (34134.0bp,86.13bp) and (34129.0bp,77.562bp)  .. (230);
  \draw (34222.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (16) ..controls (5555.9bp,125.48bp) and (5251.4bp,119.55bp)  .. (5149.2bp,92.0bp) .. controls (5126.5bp,85.876bp) and (5102.9bp,74.068bp)  .. (72);
  \draw (5220.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (19) ..controls (9025.9bp,126.13bp) and (8621.4bp,121.11bp)  .. (8487.2bp,92.0bp) .. controls (8458.0bp,85.659bp) and (8426.7bp,73.029bp)  .. (90);
  \draw (8548.2bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (3) ..controls (11491.0bp,185.16bp) and (11491.0bp,169.55bp)  .. (21);
  \draw (11567.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (26) ..controls (16130.0bp,114.59bp) and (16199.0bp,105.07bp)  .. (16225.0bp,92.0bp) .. controls (16237.0bp,86.012bp) and (16248.0bp,76.199bp)  .. (133);
  \draw (16327.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (32) ..controls (23333.0bp,122.4bp) and (23706.0bp,113.05bp)  .. (23832.0bp,92.0bp) .. controls (23874.0bp,85.063bp) and (23919.0bp,71.375bp)  .. (175);
  \draw (23976.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (29) ..controls (19879.0bp,111.93bp) and (19933.0bp,103.1bp)  .. (19979.0bp,92.0bp) .. controls (20016.0bp,83.201bp) and (20057.0bp,70.394bp)  .. (154);
  \draw (20103.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (35) ..controls (27091.0bp,107.03bp) and (27101.0bp,100.36bp)  .. (27109.0bp,92.0bp) .. controls (27115.0bp,85.554bp) and (27120.0bp,77.115bp)  .. (192);
  \draw (27190.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (29) ..controls (19770.0bp,100.22bp) and (19770.0bp,87.936bp)  .. (19770.0bp,77.0bp) .. controls (19770.0bp,74.448bp) and (19770.0bp,71.791bp)  .. (152);
  \draw (19852.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (0) ..controls (19508.0bp,273.2bp) and (19629.0bp,246.81bp)  .. (6);
  \draw (19669.0bp,258.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (24) ..controls (13526.0bp,112.21bp) and (13566.0bp,103.88bp)  .. (13600.0bp,92.0bp) .. controls (13622.0bp,84.484bp) and (13645.0bp,73.057bp)  .. (119);
  \draw (13704.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (8) ..controls (26012.0bp,193.29bp) and (25419.0bp,151.19bp)  .. (33);
  \draw (25852.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (21) ..controls (11491.0bp,98.163bp) and (11491.0bp,82.548bp)  .. (107);
  \draw (11562.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (15) ..controls (4458.8bp,107.03bp) and (4469.2bp,100.36bp)  .. (4477.2bp,92.0bp) .. controls (4483.3bp,85.554bp) and (4487.9bp,77.115bp)  .. (69);
  \draw (4503.7bp,84.5bp) node {$1/2$};
  \draw [black,->] (44) ..controls (35779.0bp,98.163bp) and (35779.0bp,82.548bp)  .. (239);
  \draw (35850.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (25) ..controls (14844.0bp,115.11bp) and (14923.0bp,105.82bp)  .. (14992.0bp,92.0bp) .. controls (15035.0bp,83.436bp) and (15082.0bp,70.024bp)  .. (127);
  \draw (15133.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (44) ..controls (35581.0bp,123.06bp) and (35376.0bp,115.18bp)  .. (35306.0bp,92.0bp) .. controls (35288.0bp,86.013bp) and (35270.0bp,74.976bp)  .. (236);
  \draw (35367.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (17) ..controls (6939.8bp,119.8bp) and (6866.9bp,111.12bp)  .. (6805.2bp,92.0bp) .. controls (6783.0bp,85.136bp) and (6759.8bp,73.527bp)  .. (81);
  \draw (6886.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (18) ..controls (7967.7bp,118.09bp) and (8077.7bp,108.54bp)  .. (8117.2bp,92.0bp) .. controls (8131.6bp,85.987bp) and (8145.3bp,75.584bp)  .. (89);
  \draw (8208.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (31) ..controls (21498.0bp,124.21bp) and (21888.0bp,116.9bp)  .. (22018.0bp,92.0bp) .. controls (22053.0bp,85.413bp) and (22090.0bp,72.334bp)  .. (165);
  \draw (22141.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (44) ..controls (36034.0bp,124.21bp) and (36424.0bp,116.9bp)  .. (36554.0bp,92.0bp) .. controls (36589.0bp,85.413bp) and (36626.0bp,72.334bp)  .. (244);
  \draw (36677.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (9) ..controls (30506.0bp,193.83bp) and (29850.0bp,150.6bp)  .. (37);
  \draw (30320.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (0) ..controls (18731.0bp,289.16bp) and (15405.0bp,228.63bp)  .. (4);
  \draw (17386.0bp,258.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (5) ..controls (18668.0bp,194.65bp) and (17913.0bp,149.78bp)  .. (27);
  \draw (18375.0bp,171.5bp) node {$1$};
  \draw [black,->] (42) ..controls (34197.0bp,112.3bp) and (34143.0bp,103.45bp)  .. (34095.0bp,92.0bp) .. controls (34059.0bp,83.354bp) and (34020.0bp,70.604bp)  .. (229);
  \draw (34111.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (19) ..controls (9442.4bp,118.48bp) and (9552.7bp,109.06bp)  .. (9592.2bp,92.0bp) .. controls (9605.9bp,86.053bp) and (9618.9bp,75.794bp)  .. (97);
  \draw (9676.2bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (31) ..controls (21341.0bp,114.44bp) and (21368.0bp,106.03bp)  .. (21389.0bp,92.0bp) .. controls (21399.0bp,85.923bp) and (21406.0bp,76.751bp)  .. (161);
  \draw (21476.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (31) ..controls (21145.0bp,114.44bp) and (21119.0bp,106.03bp)  .. (21097.0bp,92.0bp) .. controls (21088.0bp,85.923bp) and (21080.0bp,76.751bp)  .. (159);
  \draw (21168.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (25) ..controls (14842.0bp,114.59bp) and (14911.0bp,105.07bp)  .. (14937.0bp,92.0bp) .. controls (14949.0bp,86.012bp) and (14960.0bp,76.199bp)  .. (126);
  \draw (14973.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (11) ..controls (1199.7bp,121.19bp) and (1395.0bp,112.12bp)  .. (1463.2bp,92.0bp) .. controls (1484.4bp,85.733bp) and (1506.3bp,74.114bp)  .. (53);
  \draw (1569.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (11) ..controls (1163.0bp,118.64bp) and (1273.7bp,109.28bp)  .. (1313.2bp,92.0bp) .. controls (1326.8bp,86.057bp) and (1339.5bp,75.798bp)  .. (52);
  \draw (1398.2bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (44) ..controls (35877.0bp,114.44bp) and (35904.0bp,106.03bp)  .. (35925.0bp,92.0bp) .. controls (35935.0bp,85.923bp) and (35942.0bp,76.751bp)  .. (240);
  \draw (36012.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (28) ..controls (18767.0bp,119.71bp) and (18616.0bp,110.24bp)  .. (18562.0bp,92.0bp) .. controls (18544.0bp,85.852bp) and (18526.0bp,74.802bp)  .. (145);
  \draw (18633.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (32) ..controls (22855.0bp,123.82bp) and (22559.0bp,116.31bp)  .. (22459.0bp,92.0bp) .. controls (22433.0bp,85.704bp) and (22406.0bp,73.573bp)  .. (166);
  \draw (22520.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (20) ..controls (10696.0bp,97.517bp) and (10715.0bp,79.93bp)  .. (103);
  \draw (10775.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (6) ..controls (19769.0bp,185.16bp) and (19769.0bp,169.55bp)  .. (29);
  \draw (19845.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (16) ..controls (5689.3bp,114.44bp) and (5662.7bp,106.03bp)  .. (5641.2bp,92.0bp) .. controls (5631.9bp,85.923bp) and (5624.0bp,76.751bp)  .. (75);
  \draw (5712.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (37) ..controls (29355.0bp,125.15bp) and (29218.0bp,118.01bp)  .. (29103.0bp,92.0bp) .. controls (29072.0bp,85.044bp) and (29039.0bp,72.462bp)  .. (202);
  \draw (29164.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (27) ..controls (17436.0bp,116.52bp) and (17370.0bp,107.66bp)  .. (17313.0bp,92.0bp) .. controls (17287.0bp,84.628bp) and (17258.0bp,72.633bp)  .. (138);
  \draw (17384.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (27) ..controls (17306.0bp,125.14bp) and (16910.0bp,118.94bp)  .. (16777.0bp,92.0bp) .. controls (16746.0bp,85.566bp) and (16712.0bp,72.787bp)  .. (135);
  \draw (16838.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (41) ..controls (33346.0bp,115.21bp) and (33275.0bp,105.78bp)  .. (33249.0bp,92.0bp) .. controls (33238.0bp,86.082bp) and (33228.0bp,76.407bp)  .. (225);
  \draw (33320.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (39) ..controls (31770.0bp,114.59bp) and (31839.0bp,105.07bp)  .. (31865.0bp,92.0bp) .. controls (31877.0bp,86.012bp) and (31888.0bp,76.199bp)  .. (218);
  \draw (31956.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (20) ..controls (10632.0bp,97.517bp) and (10613.0bp,79.93bp)  .. (102);
  \draw (10639.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (8) ..controls (26303.0bp,185.16bp) and (26303.0bp,169.55bp)  .. (34);
  \draw (26359.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (3) ..controls (11067.0bp,197.66bp) and (9737.8bp,146.44bp)  .. (19);
  \draw (10586.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (32) ..controls (23083.0bp,98.163bp) and (23083.0bp,82.548bp)  .. (170);
  \draw (23099.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (21) ..controls (11589.0bp,114.44bp) and (11616.0bp,106.03bp)  .. (11637.0bp,92.0bp) .. controls (11647.0bp,85.923bp) and (11654.0bp,76.751bp)  .. (108);
  \draw (11714.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (37) ..controls (29297.0bp,122.9bp) and (29007.0bp,114.56bp)  .. (28908.0bp,92.0bp) .. controls (28880.0bp,85.556bp) and (28850.0bp,73.136bp)  .. (201);
  \draw (28979.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (37) ..controls (29621.0bp,114.44bp) and (29648.0bp,106.03bp)  .. (29669.0bp,92.0bp) .. controls (29679.0bp,85.923bp) and (29686.0bp,76.751bp)  .. (206);
  \draw (29767.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (29) ..controls (19671.0bp,114.7bp) and (19644.0bp,106.23bp)  .. (19623.0bp,92.0bp) .. controls (19614.0bp,85.921bp) and (19607.0bp,76.748bp)  .. (151);
  \draw (19694.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (33) ..controls (24881.0bp,122.9bp) and (24591.0bp,114.56bp)  .. (24492.0bp,92.0bp) .. controls (24464.0bp,85.556bp) and (24434.0bp,73.136bp)  .. (177);
  \draw (24563.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (16) ..controls (5787.2bp,98.163bp) and (5787.2bp,82.548bp)  .. (76);
  \draw (5858.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (26) ..controls (16051.0bp,107.03bp) and (16061.0bp,100.36bp)  .. (16069.0bp,92.0bp) .. controls (16075.0bp,85.554bp) and (16080.0bp,77.115bp)  .. (132);
  \draw (16150.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (0) ..controls (20024.0bp,297.73bp) and (23288.0bp,281.04bp)  .. (25947.0bp,266.0bp) .. controls (29355.0bp,246.73bp) and (33544.0bp,220.76bp)  .. (10);
  \draw (28546.0bp,258.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (19) ..controls (9516.6bp,125.56bp) and (9820.3bp,119.71bp)  .. (9922.2bp,92.0bp) .. controls (9944.6bp,85.906bp) and (9967.8bp,74.178bp)  .. (99);
  \draw (9973.7bp,84.5bp) node {$1/4$};
  \draw [black,->] (32) ..controls (23147.0bp,97.451bp) and (23192.0bp,76.913bp)  .. (171);
  \draw (23257.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (32) ..controls (22887.0bp,121.9bp) and (22687.0bp,113.27bp)  .. (22618.0bp,92.0bp) .. controls (22598.0bp,85.864bp) and (22578.0bp,74.52bp)  .. (167);
  \draw (22689.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (40) ..controls (32624.0bp,117.91bp) and (32733.0bp,108.31bp)  .. (32772.0bp,92.0bp) .. controls (32787.0bp,85.982bp) and (32801.0bp,75.578bp)  .. (223);
  \draw (32862.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [red,->] (30) ..controls (20268.0bp,105.71bp) and (20277.0bp,99.171bp)  .. (20285.0bp,92.0bp) .. controls (20293.0bp,84.831bp) and (20301.0bp,75.826bp)  .. (155);
  \draw (20309.0bp,84.5bp) node {$1$};
  \draw [black,->] (28) ..controls (18824.0bp,117.89bp) and (18776.0bp,109.13bp)  .. (18735.0bp,92.0bp) .. controls (18720.0bp,85.55bp) and (18705.0bp,75.06bp)  .. (146);
  \draw (18796.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (13) ..controls (3147.2bp,97.451bp) and (3102.8bp,76.913bp)  .. (61);
  \draw (3146.7bp,84.5bp) node {$1/2$};
  \draw [black,->] (27) ..controls (17722.0bp,120.21bp) and (17842.0bp,111.46bp)  .. (17944.0bp,92.0bp) .. controls (17985.0bp,84.212bp) and (18030.0bp,70.776bp)  .. (143);
  \draw (18076.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (19) ..controls (9173.2bp,112.3bp) and (9118.8bp,103.45bp)  .. (9071.2bp,92.0bp) .. controls (9035.3bp,83.354bp) and (8995.6bp,70.604bp)  .. (93);
  \draw (9086.7bp,84.5bp) node {$1/4$};
%
\end{tikzpicture}