### 2019-04-01 3-Appendices/1-Appendix/appendix1.tex
WOrking on the whole document but not yet Chapter5

%!TEX root = ../PhD_thesis__Lilian_Besson
% ******************************* Thesis Appendix A ****************************
\chapter{Mathematical Tools}
\label{app:1:mathsTools}

In this appendix, blabla.

% ----------------------------------------------------------------------------
\section{Concentration Inequalities}
\label{sec:app1:ConcentrationInequalities}

I probably need to include Hoeffding's inequality, Chernoff's method etc.

Reference: [Boucheron et al], [?]

% ----------------------------------------------------------------------------
\section{Useful lemmas}
\label{sec:app1:UsefulLemmas}

I probably need to include a bunch of useful lemmas, used in various places in the thesis, like Pinsker's inequality, etc.

% ----------------------------------------------------------------------------
\section{Results about klUCB}
\label{sec:app1:resultsklUCB}

I probably need to include some results about klUCB, mainly coming from the journal [Garivier et al], used in various places in the thesis.