### 2019-04-17 3-Appendices/1-Appendix/appendix1.tex
Cette thèse me semble interminable...

%!TEX root = ../PhD_thesis__Lilian_Besson
% ******************************* Thesis Appendix A ****************************
\chapter{Additional Mathematical Tools}
\label{app:1:mathsTools}

In this appendix, we present some mathematical tools that are used in various proofs in this thesis.

% ----------------------------------------------------------------------------
\section{Concentration Inequalities}
\label{sec:app1:ConcentrationInequalities}

\TODOL{I probably need to include Hoeffding's inequality, Chernoff's method etc?}

Reference: \cite{Boucheron2013} and maybe the original paper by Hoeffding, Chernoff etc.

% ----------------------------------------------------------------------------
\section{Useful lemmas}
\label{sec:app1:UsefulLemmas}

\TODOL{I probably need to include a bunch of useful lemmas, used in various places in the thesis, like Pinsker's inequality, etc.}

% ----------------------------------------------------------------------------
\section{Results about klUCB}
\label{sec:app1:resultsklUCB}

\TODOL{I probably need to include some results about klUCB, mainly coming from \cite{KLUCBJournal}, used in various places in the thesis?}