### 2019-04-18 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Done for now.

% % -----------------------------------------------------------------
% \section{Regret Decompositions}
% \label{app:5:regretdec}

% % -----------------------------------------------------------------
% \subsection{Proof of Lemma~\ref{lem:5:DecompositionRegret}}
% \label{proof:5:DecompositionRegret}

%     Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t):=\indic(\overline{C^j(t)})$,
%     \begin{align*}
%       R(T)
%       &= \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M Y_{A^j(t),t} \eta^j(t) \right]
%        = \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M \mu_{A^j(t)} \eta^j(t)\right]
%       \intertext{The last equality comes from the linearity of expectations, and the fact that $\E_{\mu}[Y_{k,t}] = \mu_k$ (for all $t$, from the \iid{} hypothesis), and the independence from $A^j(t)$, $\eta^j(t)$ and $Y_{k,t}$ (observed \emph{after} playing $A^j(t)$). So $\E_{\mu}[Y_{A^j(t),t} \eta^j(t)] = \sum_{k} \E_{\mu}[\mu_k \indic(A^j(t),t) \eta^j(t)] = \E_{\mu}[\mu_{A^j(t)} \eta^j(t)]$. And so}
%       R(T)
%       &= \E_{\mu}\left[ \sum_{t=1}^{T}\sum_{j \in \Mbest} \mu_j
%         - \sum_{t=1}^{T} \sum_{j=1}^{M} \mu_{A^j(t)} \eta^j(t) \right] \\
%         % &= \underbrace{\left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)}_{:= \overline{\mu}^*} (T M)
%       &= \left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)
%         - \sum_{k=1}^{K} \sum_{j=1}^{M} \mu_k \E_{\mu}\left[ T^j_k(T) \right]
%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
%       %
%       \intertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* := \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
%       %
%       &= \sum_{k=1}^{K} \sum_{j=1}^{M} (\overline{\mu}^* - \mu_k) \E_{\mu}\left[ T^j_k(T) \right]
%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
%       %
%       \intertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $\{1,\dots,K\}$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
%       %
%       &= \sum_{k \in \Mbest\setminus\{M\}} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
%         + \overline{\Delta_{M^*}} \E_{\mu}\left[ T_{M^*}(T) \right] \\
%         &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
%       %
%       \intertext{But for $k = M^*$, $T_{M^*}(T) = T M^* - \sum\limits_{k \in \Mbest\setminus\{M\}} \E_{\mu}\left[ T_k(T) \right] - \sum\limits_{k \in \Mworst} \E_{\mu}\left[ T_k(T) \right]$, so by recombining the terms, we obtain,}
%       %
%       &= \sum_{k \in \Mbest\setminus\{M\}} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
%         + \overline{\Delta_{M^*}} T M^* \\
%         &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
%     \end{align*}
%     %
%     The term $\overline{\Delta_k} - \overline{\Delta_{M^*}}$ simplifies to $\mu_{M^*} - \mu_k$, and so $\overline{\Delta_{M^*}} = \frac{1}{M} \sum_{k=1}^{M} \mu_k - \mu_{M^*}$ by definition of $\overline{\mu}^*$. And for $k=M^*$, $\mu_{M^*} - \mu_k = 0$, so the first sum can be written for $k = 1,\dots,M$ only, so
%     %
%     % \vspace*{5pt}
%     \begin{align*}
%       R(T)
%       &= \sum_{k \in \Mbest} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
%         + \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) T \\
%         &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right]
%     \end{align*}
%     And so we obtain the decomposition with three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3}.
%     \begin{align*}
%     R(T)
%       &= \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) \left(T - \E_{\mu}\left[ T_k(T) \right]\right)
%         + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
%     \end{align*}
%     % Which is exactly the decomposition we wanted to prove.

% % -----------------------------------------------------------------
% \subsection{Proof of Lemma~\ref{lem:5:1stLowerBound}}
% \label{proof:5:1stLowerBound}

%   Note that term \ref{eq:5:term3} is clearly lower bounded by $0$
%   but it is not obvious for \ref{eq:5:term2} as there is no reason for $T_k(T)$ to be upper bounded by $T$.
%   %
%   Let $T_k^{!}(T) := \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
%   where the notation $\exists!$ stands for ``there exists a unique''.
%   Then $T_k(T) = \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(A^j(t) = k)$ can be decomposed as
%   \begin{equation*}
%     T_k(T) = \sum_{t=1}^{T} \indic(\exists! j, A^j(t) = k) + \sum_{t=1}^{T} \sum_{j=1}^{M} c_{k,t} \indic(A^j(t) = k)
%     = T_k^{!}(T) + C_k(T).
%   \end{equation*}
%   By focusing on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T(\boldsymbol{\mu}, M, \rho)$ from Lemma~\ref{lem:5:DecompositionRegret}, we have
%   \begin{align*}
%     \ref{eq:5:term2} + \ref{eq:5:term3} &=
%     \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
%     + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)] \\
%     & \;\;\;\;\;\; + \sum_{k=1}^{M} \mu_k \E_{\mu}[C_k(T)]
%     - \sum_{k \in \Mbest} \mu_k \E_{\mu}[C_k(T)] \\
%     &=
%     \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
%     + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)]
%     + \sum_{k \in \Mworst} \mu_k \E_{\mu}[C_k(T)] \\
%     &=
%     \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
%     + \sum_{k=1}^{M} \min(\mu_M^*, \mu_k) \E_{\mu}[C_k(T)].
%   \end{align*}
%   And now both terms are non-negative, as $T_k^!(T) \leq T$, $\min(\mu_M^*, \mu_k)\geq 0$, and $C_k(T) \geq 0$, so $\ref{eq:5:term2} + \ref{eq:5:term3} \geq 0$
%   which proves that $R_T(\boldsymbol{\mu}, M, \rho) \geq \ref{eq:5:term1}$, as wanted.

% % -----------------------------------------------------------------
% \subsection{Proof of Lemma~\ref{lem:5:1stUpperBound}}
% \label{proof:5:1stUpperBound}

% Recall that we want to upper bound
% $ (b) : = \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[T_k(T)]\right)$.
% First, we observe that, for all $k\in \Mbest$,
% \begin{eqnarray*}
%   T - \E_{\mu}[T_k(T)] & \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] \\
%   & = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right],
% \end{eqnarray*}
% where we denote by $\widehat{S}_t = \{A^j(t), j \in \{1,\dots,M\}\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
% \begin{eqnarray*}
%  (b) & \leq & (\mu_1 - \mu_{M^*})  \sum_{k \in \Mbest} \left(T - \E_{\mu}[T_k(T)]\right) \leq  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[\sum_{k \in \Mbest} \sum_{t = 1}^T \indic(k \notin \widehat{S}_t)\right] \\
%  & = &  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right].
% \end{eqnarray*}
% The quantity $\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)$ counts the number of optimal arms that have not been selected at time $t$. For each mis-selection of an optimal arm, there either exists a sub-optimal arm that has been selected, or an arm in $\Mbest$ on which a collision occurs. Hence
% \[\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t) = \sum_{k \in \Mbest}\indic(C_k(t)) + \sum_{k \in \Mworst} \indic(\exists j : A^j(t) = k),\]
% which yields
% \[\E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right] \leq \sum_{k \in \Mbest}\E_{\mu}\left[\cC_k(T)\right] + \sum_{k \in \Mworst} \E_{\mu}\left[T_k(T)\right]\]
% and Lemma~\ref{lem:5:1stUpperBound} follows.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\subsection{Lower bound: proof of Theorem~\ref{thm:5:BetterLowerBound}}
\label{proof:5:BetterLowerBound}

The lower bound that we present relies on the following \emph{change-of-distribution} lemma that we prove in the next section, following recent arguments from \cite{Garivier16TrueShape} that have to be adapted to incorporate the collision information.

\begin{lemma}\label{lem:5:CD} Under observation model \modelun{} and \modeldeux, for every event $A$ that is $\cF_{T}^{j}$-measurable, considering two multi-player bandit models denoted by $\bm{\mu}$ and $\bm{\lambda}$ respectively, it holds that
\begin{equation}
  \sum_{k=1}^{K}\E_{\bm{\mu}}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right).
\end{equation}
\end{lemma}

Let $k$ be a sub-optimal arm under $\bm{\mu}$,
fix $\varepsilon \in \left(0, \mu_{M-1}^* - \mu_M^*\right)$,
% We define the bandit instance $\bm{\lambda}$ such that
and let $\bm{\lambda}$ be the bandit instance such that
\[\left\{\begin{array}{ccl}
          \lambda_\ell & = & \mu_\ell \ \ \  \text{for all } \ell \neq k, \\
          \lambda_k & = & \mu_{M^*} + \varepsilon.
         \end{array}
\right.\]
Clearly, $\bm{\lambda}\in\cP_M$ also,
and the set of $M$ best arms under $\bm{\mu}$ and $\bm{\lambda}$ differ by one arm: if \Mbest$_{\bm{\mu}} = \{1^*,\dots,M^*\}$ then $\Mbest_{\bm{\lambda}} = \{1^*,\dots,(M-1)^*,k\}$.
Thus, one expects the ($\cF^j_T$-mesurable) event
\[A_T = \left(T_k^j(T) > \frac{T}{2M}\right)\]
to have a small probability under $\bm{\mu}$ (under which $k$ is sub-optimal) and a large probability under $\bm{\lambda}$ (under which $k$ is one of the optimal arms, and is likely to be drawn a lot).

Applying the inequality in Lemma~\ref{lem:5:CD}, and noting that the sum in the left-hand side reduces to on term as there is a single arm whose distribution is changed, one obtains
\begin{eqnarray*}
  \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \kl \left(\Pr_{\bm{\mu}}(A_T),\Pr_{\bm{\lambda}}(A_T)\right),\nonumber \\
  \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - \Pr_{\bm{\mu}}(A_T)\right) \log \left(\frac{1}{\Pr_{\bm{\lambda}}(\overline{A_T})}\right) - \log(2),
\end{eqnarray*}
using the fact that the binary KL-divergence satisfies $\kl(x,y) = \kl(1-x,1-y)$ as well as the inequality  $\kl(x, y) \geq x \log\left(1/y\right) - \log(2)$, proved by \cite{Garivier16TrueShape}.
Now, using Markov inequality yields
\begin{align*}
  \Pr_{\bm{\mu}}\left(A_T\right)
    & \leq 2M\frac{\E_{\bm{\mu}}\left[T^j_k(T)\right]}{T} =: x_T, \\
  \Pr_{\bm{\lambda}}\left(\overline{A_T}\right)
    & = \Pr_{\bm{\lambda}}\left(\frac{T}{M} - T^j_k(T) > \frac{T}{2M}\right) \leq 2M\frac{\E_\bm{\lambda}\left[ \frac{T}{M} - T^j_k(T)\right]}{T} =: \frac{y_T}{T},
\end{align*}
which defines two sequences $x_T$ and $y_T$, such that
\begin{eqnarray}\label{eq:5:CqCD}
  \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - x_T\right) \log \left(\frac{T}{y_T}\right) - \log(2).
\end{eqnarray}
The strong uniform efficiency assumption  (see Definition~\ref{def:5:DecentralizedUniformEfficiency}) further tells us that $x_T \rightarrow 0$ (as $\E_{\bm{\mu}}[T_k^j(T)] = o(T^\alpha)$ for all $\alpha$) and $y_T = o(T^\alpha)$ when $T \to \infty$, for all $\alpha \in (0,1)$.
As a consequence, observe that $\log(y_T)/\log(T) \rightarrow 0$ when $T$ tends to infinity and
\[
  \frac{\left(1 - x_T\right) \log \left({T}/{y_T}\right) - \log(2)}{\log(T)} = 1-x_T - (1-x_T) \frac{\log(y_T)}{\log(T)} - \frac{\log(2)}{\log(T)}
\]
tends to one when $T$ tends to infinity.
From Equation~\eqref{eq:5:CqCD}, this yields
\begin{equation}\label{eq:5:CqCD2}
  \liminf_{T\to\infty} \frac{\E_{\bm{\mu}}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k,\mu_{M^*}+\varepsilon)},
\end{equation}
for all $\varepsilon \in \left(0, \mu_{M-1}^* - \mu_M^*\right)$.
Letting $\varepsilon$ go to zero gives the conclusion (as $\kl$ is continuous).

% -----------------------------------------------------------------

\subsubsection{Proof of Lemma~\ref{lem:5:CD}}

Under observation model \modelun{} and \modeldeux, the strategy $A^j(t)$ decides which arm to play based on the information contained in $\cO_{t-1}$, where $\cO_0 = U_0$ and
\[\forall t > 0, \ \ \cO_t = (U_0,Y_1,C_1,U_1,\dots,Y_t,C_t,U_t)\]
where $Y_{t} := Y_{A^j(t-1),t}$ denotes the sensing information, $C_t := C^j(t)$ denotes the collision information (not always completely exploited under observation model \modeldeux) and $U_t$ denotes some external source of randomness\footnote{For instance, \MCTopM, \RandTopM{} and \rhoRand{} draws from a uniform variable in $\{1,\dots,M\}$ for new ranks or arms.} useful to select $A^j(t)$.
Formally, one can say that $A^j(t)$ is $\sigma(\cO_{t-1})$ measurable, if $\sigma(\cO_t)$ denotes the sigma-Algebra generated by the observations $\cO_t$, and as $\cF^j(t) \subseteq \sigma(\cO_{t})$, with an equality under observation model \modelun.

Under two bandit models $\bm{\mu}$ and $\bm{\lambda}$, we let $\Pr_{\bm{\mu}}^{\cO_t}$ (resp. $\Pr_{\bm{\lambda}}^{\cO_t}$) be the distribution of the observations under model $\bm{\mu}$ (resp. $\bm{\lambda}$), given a fixed algorithm.
Using the exact same technique as \cite{Garivier16TrueShape}, called the ``contraction of entropy'' principle, one can establish that for any event $A$ that is $\sigma(\cO_t)$-measurable\footnote{In the work of \cite{Garivier16TrueShape}, the statement is more general and the probability of an event $A$ is replaced by the expectation of any $\cF_T$-measurable random variable $Z$ bounded in $[0,1]$.},
\[\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right) \geq \kl\left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right).\]

The next step is to relate the complicated KL-divergence $\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right)$ to the number of arm selections.
Proceeding similarly as \cite{Garivier16TrueShape}, one can write, using the chain rule for KL-divergence, that
\begin{equation}\label{eq:5:Iterative}\KL\left(\Pr_{\bm{\mu}}^{\cO_{t}},\Pr_{\bm{\lambda}}^{\cO_{t}}\right) =  \KL\left(\Pr_{\bm{\mu}}^{\cO_{t-1}},\Pr_{\bm{\lambda}}^{\cO_{t-1}}\right) + \KL\left(\Pr_{\bm{\mu}}^{Y_{t},C_t,U_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t},C_t,U_t | \cO_{t-1}}\right).\end{equation}
Now observe that conditionally to $\cO_{t-1}$, $U_t$, $Y_t$ and $C_t$ are independent, as once the selected arm is known, the value of the sensing $Y_t$ does not influence the other players selecting that arm, and $U_t$ is some exogenous randomness.
Using further that the distribution of $U_t$ is the same under $\bm{\mu}$ and $\bm{\lambda}$, one obtains
\begin{equation}\KL\left(\Pr_{\bm{\mu}}^{Y_{t},C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t},C_t | \cO_{t-1}}\right) =
 \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right) + \KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right).\label{eq:5:DecKL}
\end{equation}

The first term in \eqref{eq:5:DecKL} can be rewritten using the same argument as \cite{Garivier16TrueShape}, that relies on the fact that conditionally to $\cO_{t-1}$, $Y_t$ is a Bernoulli distribution with mean $\mu_{A^j(t)}$ under the instance $\bm{\mu}$ and $\lambda_{A^j(t)}$ under the instance $\bm{\lambda}$:
\begin{eqnarray*}
\KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right)  & = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right) | \cO_{t-1} \right]\right] \\
& = & \E_{\bm{\mu}} \left[\kl \left(\mu}_{A^j(t)},\lambda_{A^j(t)}\right)\right] \\
& = & \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu}_{k},\lambda_{k}\right)\right],
\end{eqnarray*}
We now show that second term in \eqref{eq:5:DecKL} is zero:
\begin{eqnarray*}
\KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right)
& = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{C_t| \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t| \cO_{t-1}}\right) | \cO_{t-1} \right]\right] \\
& = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{C_t| \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t| \cO_{t-1}}\right) | \bigcup_{j' \neq j} \cO_{t-1}^{j'}\right]| \cO_{t-1} \right]\right], \\
\end{eqnarray*}
where $\cO_{t-1}^{j'}$ denote the information available to player $j' \neq j$.
\textcolor{red}{By hypothesis}, the ``collision information term'' is assumed to be a $\smallO{\log(T)}$,
and thus $\KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right) = \smallO{\log(T)}$.

Putting things together, we showed that
\[\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right) = \KL\left(\Pr_{\bm{\mu}}^{\cO_{t-1}},\Pr_{\bm{\lambda}}^{\cO_{t-1}}\right) + \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu}_{k},\lambda_{k}\right)\right] + \smallO{\log(T)}.\]
Iterating this equality and using that $\KL\left(\Pr_{\bm{\mu}}^{\cO_{0}},\Pr_{\bm{\lambda}}^{\cO_{0}}\right)=0$ yields that
\[\sum_{k=1}^{K}\E_{\bm{\mu}}\left[T_k^j(T)\right]\kl(\mu}_k,\lambda_k) \geq \kl \left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right) + \smallO{\log(T)},\]
for all $A \in \sigma(\cO_T)$, in particular for all $A \in \cF^j_T$.

% -----------------------------------------------------------------
% -----------------------------------------------------------------

% \section{Proofs Elements Related to Regret Upper Bounds}
% \label{proof:5:RegretUpperBounds}

% This Appendix includes the main proofs, missing from the content of the article,
% that yield the regret upper bound.
% We start by controlling the sub-optimal draws when the \klUCB{} indices are used (instead of \UCB),
% with any of our proposed algorithms (\MCTopM, \RandTopM) or \rhoRand{}. Then we focus on controlling collisions for \MCTopM-\klUCB.

% \subsection{Control of the Sub-optimal Draws for \klUCB: Proof of Lemma~\ref{lem:5:SubOptimalSelections}}
% \label{proof:5:SubOptimalSelections}

% %
% %     Recall $g_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
% %     As only user $j$ is considered here, the superscript $j$ is dropped.
% %
% Fix $k\in\Mworst$ and a player $j \in \{1,\dots,M\}$.
% The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
% \begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : g_m^j(t) < g_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
% Indeed, for the three algorithms, an arm selected at time $t+1$ belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
% If the sub-optimal arm $k$ is selected at time $t$, it implies that $k \in \TopM(t)$, and, because there are $M$ arms in both \Mbest{} and $\TopM(t)$, one of the arms in \Mbest{} must be excluded from $\TopM(t)$.
% In particular, the index of arm $k$ must be larger than the index of this particular arm $m$.

% Using \eqref{eq:5:KeyInclusion}, one can then upper bound the number of selections of arm $k$ by user $j$ up to round $T$ as
% \begin{align*}
% \E_{\mu}[T_k^j(T)]
% &= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
% = \sum_{t=1}^T \Pr\left( A^j(t) = k \right).\\
% %&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
% &= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
% %
% \intertext{Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition}
% \E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
% &\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr\left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
% &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr\left(g_{m^*}(t) < \mu_m^* \right) \\
% %
% &\leq
% \sum_{t=1}^T
%   \Pr\left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
% +
% \sum_{m=1}^M
%   \sum_{t=1}^T \Pr\left( g_{m^*}(t) < \mu_m^* \right),
% \end{align*}
% where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in \{1, \dots,M\}$.

% Now each of the two terms in the right hand side can directly be upper bounded using tools developed by \cite{KLUCBJournal} for the analysis of kl-UCB.
% The rightmost term can be controlled using Lemma~\ref{lem:5:Fact1KLUCB} below that relies on a self-normalized deviation inequality, whose proof exactly follows from the proof of Fact 1 in Appendix A of \cite{KLUCBJournal}.
% The leftmost term can be controlled using Lemma~\ref{lem:5:Fact2KLUCB} stated below, that is a direct consequence of the proof of Fact 2 in Appendix A of \cite{KLUCBJournal}.

% \begin{lemma}\label{lem:5:Fact1KLUCB}
%     For any arm $k$, if $g_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
%     \begin{equation}
%       \sum_{t=1}^T \Pr\left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
%     \end{equation}
% \end{lemma}

% Denote $\kl'(x,y)$ the derivative of the function $x \mapsto \kl(x,y)$ (for any fixed $y\neq 0, 1$).

% \begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
% \begin{align*}
%   &\sum_{t=1}^T\Pr\left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
%   + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.\end{align*}
% \end{lemma}

% Putting things together, one obtains the non-asymptotic upper bound
% \begin{align}\label{eq:5:UBprecise}
%   \E_{\mu}\left[T_k^j(T)\right]
%   & \leq \frac{\log(T) + 3 \log\log(T)}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log\log(T)}  \nonumber\\
%   & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4Me \log\log(T) + 3M+1,
% \end{align}
% which yields Lemma~\ref{lem:5:SubOptimalSelections},
% with explicit constants $C_{\boldsymbol{\mu}}$ and $D_{\boldsymbol{\mu}}$.

% % --------------------------------------------------------------------------------
% \subsection{Proof of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}}\label{proof:5:elementaryLemma_RandTopM_MCTopM}

% Using the behavior of the algorithm when the current arm leaves the set $\TopM$ (Line 4), one has
% \begin{align*}
%   & \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
%   & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
%   & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
% \end{align*}

% Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
% \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
% The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}. To control the second term, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
% Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
% and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
% \begin{align}
%   & \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)
%   =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
%   & \hspace{1cm} \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
%   & \hspace{1cm}  \leq \sum_{s=1}^T \bP\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
% \end{align}
% where the last inequality uses that for all $s$, \[\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} = \sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t) = s + 1\right)} \leq 1.\]
% From \eqref{eq:5:FromHere}, the same upper bound as that of Lemma~\ref{lem:5:Fact2KLUCB} can be obtained using the tools from \cite{KLUCBJournal}, which proves that for $T\to\infty$,
% \[T_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]

% If $\mu_k > \mu_{k'}$, we rather use that
% \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \bP\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
% and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
% \[T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T)).\]
% Summing over $k'$ yields the result.

%%%%%%%%%%%%ù PREVIOUS VERSION
% For any arms $k, k' \in \{1, \dots, K\}$,
% let $\phi_{k,k'}^j(t)$ be the event
% \begin{equation*}
%   % \phi_{k}^j(t) :=
%   \left(
%     % \overline{C^j(t-1)} \cap
%     A^j(t-1) = k \in \TopM(t-1),
%     k \notin \TopM(t),
%     C^j(t),
%     A^j(t) = k' \in \TopM(t) \cap \{ m : g_m^j(t-2) \leq g_k^j(t-2) \}
%   %
%   \right).
% \end{equation*}
%
%
% If $k$ or $k' \in \Mworst$, the desired result directly comes from
% Lemma~\ref{lem:5:SubOptimalSelections} controlling the selections of worst arms,
% as $\phi_{k,k'}^j(t)$ implies $A^j(t) = k'$ and $A^j(t-1) = k$,
% and so $\sum_{t=1}^T \Pr(\phi_{k,k'}^j(t)) \leq \min(\E_{\mu}[T_k^j(T-1)], \E_{\mu}[T_{k'}^j(T-1)])$,
% and at least one term is $\bigO{\log T}$ from Lemma~\ref{lem:5:SubOptimalSelections}
% and both are always $\bigO{T}$,
% so $\sum_{t=1}^T \phi_{k,k'}^j(t) = \bigO{\log T}$.
%
% If one of the two arms $k, k'$ is not sampled enough, \ie, if $\E_{\mu}[T_m^j(T)]\not\to \infty$ when $T\to\infty$ for $m=k$ or $m=k'$, then these sums are $\sum_{t=1}^T \Pr(\phi_{k,k'}^j(t)) = \bigO{1} = \smallO{\log T}$ and we are done.
%
% Now assume that $k \in \Mbest$ and $k' \in \Mbest$,
% and assume that both arms are sampled enough.
% %
% This event $\phi_{k,k'}^j(t)$ we focus on implies both
% $g_{k'}^j(t-1) \leq g_k^j(t-1), A^j(t-1)=k$
% and
% $g_k^j(t) \leq g_{k'}^j(t), A^j(t)=k'$.
% %
% In particular,
% if $\mu_k \geq \mu_{k'}$,
% $\phi_{k,k'}^j(t)$ implies $g_k^j(t) \leq g_{k'}^j(t)$,
% and
% if $\mu_k < \mu_{k'}$, then
% $\phi_{k,k'}^j(t)$ implies $g_{k'}^j(t-1) \leq g_k^j(t-1)$,
% both with $A^j(t-1)=k, A^j(t)=k'$.
% Both cases are dealt with exactly the same arguments,
% so let focus on the first one.
%
% Assume $\mu_k \geq \mu_{k'}$.
% Then, essentially $g_k^j(t) \leq g_{k'}^j(t)$
% is unlikely (\ie, the sum of probabilities is $\bigO{\log T}$)
% as the indices $g_m^j(t)$ tends to $\mu_m$ if arm $m$ is sampled enough (\ie, $\E_{\mu}[T_m^j(T)] \to \infty$).
% \begin{align}
%     \sum_{t=1}^{T} \Pr(\phi_{k,k'}^j(t))
%     &\leq
%     \sum_{t=1}^{T} \Pr\left(g_k^j(t) \leq g_{k'}^j(t), A^j(t-1)=k, A^j(t)=k'\right)
%     \intertext{Either $g_{k'}^j(t) > \mu_k$ or $g_{k'}^j(t) \leq \mu_k$, and by introducing the two cases,}
%     &\leq
%     \sum_{t=1}^{T} \Pr\left(g_k^j(t) \leq g_{k'}^j(t) \leq \mu_k, A^j(t-1)=k, A^j(t)=k'\right) \notag \\
%     &\;\;+ \sum_{t=1}^{T} \Pr\left(g_k^j(t) \leq g_{k'}^j(t), g_{k'}^j(t) > \mu_k, A^j(t-1)=k, A^j(t)=k'\right) \notag
%     \intertext{For the first term, $g_k^j(t) \leq \mu_k$ is highly unlikely (\ie, the sum of probabilities is $\smallO{\log T}$) as the indices ${(g_k^j)}_k$ are Upper Confidence Bounds (if arm $k$ is sampled enough). For the second term, $g_{k'}^j(t) > \mu_k$ is unlikely if arm $k'$ is sampled enough. More formally,}
%     \sum_{t=1}^{T} \Pr(\phi_{k,k'}^j(t))
%     &\leq
%     \underbrace{\sum_{t=1}^{T} \Pr\left( g_k^j(t) \leq \mu_k \right)}_{(i)}
%     + \underbrace{\sum_{t=1}^{T} \Pr\left( g_{k'}^j(t) > \mu_k, A^j(t)=k' \right)}_{(ii)}.
% \end{align}
% But then as both arms are sampled enough, we can use both Lemma~\ref{lem:5:Fact1KLUCB} to directly obtain that the first term $(i)$ is a $\smallO{\log T}$,
% and Lemma~\ref{lem:5:Fact2KLUCB} to obtain that the second  $(ii)$ is a $\bigO{\log T}$.
%
%
% Exactly the same arguments work for the other case $\mu_k < \mu_{k'}$, by considering
% whether $g_k^j(t-1) > \mu_{k'}$ (which is also ulikely),
% or whether $g_k^j(t-1) \leq \mu_k$ which implies $g_{k'}^j(t-1) \leq \mu_{k'}$ (and which is also highly unlikely), and with Lemmas~\ref{lem:5:Fact1KLUCB}, \ref{lem:5:Fact2KLUCB},
% we obtain that
% $\sum_{t=1}^{T} \Pr(\phi_{k,k'}^j(t)) = \bigO{\log T}$ when $T \to \infty$
% similarly.
%
% Summing up for all $k' \neq k$, splitting between $\mu_{k'} < \mu_k$ or $\mu_{k'} > \mu_k$,
% and using the precise constants from Lemmas~\ref{lem:5:SubOptimalSelections}, \ref{lem:5:Fact1KLUCB}, \ref{lem:5:Fact2KLUCB} concludes.
% \vspace*{-20pt}  % XXX remove if problem

% % --------------------------------------------------------------------------------
% \subsection{Controlling Collisions for \MCTopM: Proof of Lemma~\ref{lem:5:collisionsMCTopM}}
% \label{proof:5:collisionsMCTopM}

% A key feature of both the \RandTopM{} and \MCTopM{} algorithms is Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, that states that the probability of switching from some arm because this arm leaves $\TopM(t)$ is small. Its proof is postponed to the end of this section.

% %
% Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below provides a schematic representation of the execution of the \MCTopM{} algorithm, that has to be exploited in order to properly control the number of collisions.
% %
% The sketch of the proof is the following: by focusing only on collisions in the ``not fixed'' state, bounding the number of transitions $(2)$ and $(3)$ is enough.
% Then, we show that both the number of transitions $(3)$ and $(5)$ are small: as a consequence of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, the average number of these transitions is $\bigO{\log T}$.
% Finally, we use that the length of a sequence of consecutive transitions $(2)$ is also small (on average smaller than $M$), and except for possibly the first one, starting a new sequence implies a previous transition $(3)$ or $(5)$ to arrive in the state ``not fixed''. This gives a logarithmic number of transitions $(2)$ and $(3)$, and so gives $\E_{\mu}[\sum_k\cC_k(T)] = \bigO{\log T}$,
% with explicit constants depending on $\boldsymbol{\mu}$ and $M$.

% \begin{figure}[h!]
%   \resizebox{0.98\textwidth}{!}{
%   \begin{tikzpicture}[>=latex',line join=bevel,scale=5]
%       %
%       \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
%       \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
%       \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
%       %
%       \draw [black,->] (start) -> (notfixed.20);
%       \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
%       \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
%       \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
%       \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
%       \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
%       %
%   \end{tikzpicture}
%   }
%   \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
%   Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
%   \label{fig:5:StateMachineAlgorithm_MCTopM}
% \end{figure}

% % Explanation of the figure

% As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
% Formally, $s^j(0)$ is false, and $s^j(t+1)$ is defined inductively from $s^j(t)$ as
% % following the algorithm of \MCTopM:
% \begin{equation}
%     s^j(t+1) =
%     \left( s^j(t) \cup \left( \overline{s^j(t)} \cap \overline{C^j(t)} \right) \right)
%     \cap \left( A^j(t) \in \TopM(t) \right).
% \end{equation}

% For the sake of clarity, we now explain Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in words. At step $t$, if player $j$ is not fixed ($\overline{s^j(t)}$), she can have three behaviors when executing \MCTopM.
% She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
% or she stays in state $\overline{s^j(t)}$,
% with two cases.
% Either she sampled $A^j(t+1)$ uniformly
% from $\TopM(t) \cap \{ m : g_m^j(t) \leq g_k^j(t) \}$
% with transition $(3)$,
% in case of collision and if $A^j(t+1) \in \TopM(t)$,
% or she sampled $A^j(t+1)$ uniformly
% from $\TopM(t)$ with transition $(2)$,
% if $A^j(t+1) \notin \TopM(t)$.
% In particular, note that if $\overline{C^j(t)}$, transition $(3)$ is executed and not $(2)$.
% %
% % The case $(2)$ happens when there is no collision at time $t-1$ but a collision at time $t$,
% % and conversely case $(3)$ happens when the player was not fixed on her arm at time $t-1$ (\ie, $\overline{s^j(t-1)}$)
% % and experienced a collision at time $t-1$
% Transition $(3)$ is a uniform sampling from $\TopM(t)$ (the ``Musical Chair'' step).

% For player $j$ and round $t$, we now introduce a few events that are useful in the proof. First, for every $x=1,2,3,4,5$, we denote $I_x^j(t)$ the event that a transition of type $(x)$ occurs for player $j$ after the first $t$ observations (\ie, between round $t$ and round $t+1$, to decide $A^j(t+1)$).
% Formally they are defined by
% \begin{align*}
%   I_1^j(t) &:= \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
%   % go sitted
%   I_2^j(t) &:= \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
%   \ \ \ &\text{and} \ \ \
%   I_3(t) &:= \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
%   % if sitted
%   I_3(t) &:= \left(s^j(t),A^j(t) \in \TopM(t)\right),
%   \ \ \ &\text{and} \ \ \
%   I_5(t) &:= \left(s^j(t),A^j(t) \notin \TopM(t)\right).
% \end{align*}
% %
% Then, we introduce $\widetilde{C^j}(t)$ as the event that a collision occurs for player $j$ at round $t$ if she is not yet fixed on her arm, that is
% \begin{equation}
%     \widetilde{C^j}(t) := \left(C^j(t), \overline{s^j(t)}\right).
% \end{equation}

% A key observation is that $C^j(t)$ implies $\bigcup_{j'=1}^M \widetilde{C^{j'}}(t)$, as a collision necessarily involves at least one player not yet fixed on her arm ($\overline{s^{j'(t)}}$).
% Otherwise, if they are all fixed, \ie, for all $j$, $s^j(t)$, then by definition of $s^j(t)$, none of the player changed their arm from $t-1$ to $t$, and none experienced any collision at time $t-1$ so by induction there is no collision at time $t$.
% %
% Thus, $\sum_{j=1}^M \Pr(C^j(t))$ can be upper bounded by $M \sum_{j=1}^M \Pr(\widetilde{C^j}(t))$ (union bound),
% and it follows that if $\cC(T) := \sum_{k=1}^K \cC_k(T)$ then
% \[\E_{\mu}[\cC(T)] \leq M \sum_{j=1}^M \sum_{t=1}^T \Pr(\widetilde{C^j}(t)).\]
% We can further  observe that $\widetilde{C^j}(t)$ implies a transition $(2)$ or $(3)$, as a transition $(1)$ cannot happen in case of collision. Thus another union bound gives
% %
% \begin{align}
%   \sum_{t=1}^T \Pr(\widetilde{C^j}(t))
%   &\leq \sum_{t=1}^T \Pr(I_2^j(t))   + \sum_{t=1}^T \Pr(I_3^j(t)).\label{eq:5:UBTilde}
% \end{align}
% In the rest of the proof we focus on bounding the number of transitions $(2)$ and $(3)$.

% Let $N_x^j(T)$ be the random variable denoting the number of transitions of type $(x)$.
% Neglecting the event $\overline{s^j(t)}$ for $x=3$ and $s^j(t)$ for $x=5$, one has
% \begin{equation}
%     \E_{\mu}[N_x^j(t)]
%     = \sum_{t=1}^T \Pr(I_x^j(t))
%     \leq \sum_{t=1}^T \Pr \left(A^j(t) \notin \TopM(t)\right)
%     \leq \sum_{t=1}^T \sum_{k=1}^K \Pr \left(A^j(t)=k, k \notin \TopM(t)\right),
% \end{equation}
% which is $\bigO{\log T}$ (with known constants) by Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}. In particular, this controls the second term in the right hand side of \eqref{eq:5:UBTilde}.

% To control the first term $\sum_{t=1}^T \Pr(I_2^j(t))$
% we introduce three sequences of random variables,
% the starting times $(\theta_i)_{i \geq 1}$
% and the ending times $(\tau_i)_{i \geq 1}$
% (possibly larger than $T$),
% of sequences during which $I_2(s)$ is true for all $s=\theta_i,\dots,\tau_i-1$ but not before and after,
% that is
% $\forall i \in \{1,\dots,n(T)\},
% \overline{I_2^j(\theta_i - 1)}
% \cap \bigcap_{t=\theta_i}^{\tau_i-1} I_2^j(t)
% \cap \overline{I_2^j(\tau_i)}
% $
% with $n(T)$ the number of such sequences,
% \ie,
% $n(T) := \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
% (or $0$ if $\theta_1$ does not exist).
% %
% If $\theta_i = 1$, the first sequence does not have term $\overline{I_2^j(\theta_i - 1)}$.

% Now we can decompose the sum on $t=1,\dots,T$ with the use of consecutive sequences,
% \begin{align}
%   \E_{\mu}[N_2^j(t)]
%   &= \E_{\mu}\left[ \sum_{t=1}^T \indic{\left(I_2^j(t)\right)} \right]
%   =
%   \E_{\mu}\left[ \sum_{i=1}^{n(T)} \left( \sum_{t=\theta_i}^{\tau_i - 1} 1 + \sum_{t=\tau_i}^{\theta_{i+1} - 1} 0 \right) \right]
%   =
%   \E_{\mu}\left[ \sum_{i=1}^{n(T)} (\tau_i - \theta_i) \right].
%   \intertext{Both $n(T)$ and $\tau_i - \theta_i \geq 0$ have finite averages for any $i$ (as $\tau_i - \theta_i \leq T$), and $n(T)$ is a \emph{stopping time} with respect to the past events (that is, $\cF^j_T$), and so we can obtain}
%   % so we can use Wald's Lemma \citep{Wald45}
%   &\leq \E_{\mu}\left[ n(T) \right] \times \max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]
%   = (\alpha) \times (\beta).
% \end{align}

% $(\alpha)$ To control $\E_{\mu}[n(T)]$, we can observe that
% the number of sequences $n(T)$ is smaller than $1$ plus the number of times when \emph{a sequence begins} ($1$ plus because maybe the game starts in a sequence).
% %
% And beginning a sequence
% % \footnote{Similarly, one could focus on the number of times \emph{a sequence finishes}, and $I_2^j(\theta_i) \cap \overline{I_2^j(\theta_i+1)}$ implies a transition of type $(3)$ or $(1)$.
% % The key is that after a transition $(1)$, a new sequence of transitions $(2)$ is possible only after a transition $(5)$, which total number is also controlled as $\bigO{\log T}$.}
% at time $\theta_i$ implies
% $\overline{I_2^j(\theta_i-1)} \cap I_2^j(\theta_i)$,
% which implies a transition of type $(3)$ or $(5)$ at time $\theta_i - 1$, as player j is in state ``not fixed'' at time $\theta_i$ (transitions $(1)$ and $(4)$ are impossible).
% %
% As stated above, $\E_{\mu}[N_x^j(T)] = \bigO{\log T}$ for both $x=3$ and $x=5$,
% and so $\E_{\mu}[n(T)] = \bigO{\log T}$ also.

% $(\beta)$ To control $\E_{\mu}\left[ \tau_i - \theta_i \right]$,
% a simple argument can be used.
% $\bigcup_{t=\theta_i}^{\tau_i-1} I_2^j(t)$
% implies $C^j(t)$ for $\tau_i - \theta_i$ consecutive times.
% %
% The very structure of \RandTopM{} gives that in this sequence of transitions $(2)$,
% the successive collisions (\ie, $C^j(t-1) \cap C^j(t)$)
% implies that each new arm $A^j(t+1)$ for $t \in \{\theta_i, \tau_i-1\}$ is selected uniformly from
% $\TopM(t+1)$,
% a set of size $M$ with at least one available arm.
% %
% Indeed, as there is $M-1$ other players, at time $t+1$ \emph{at least} one arm in $\TopM(t+1)$ is not selected by any player $k'\neq k$,
% and so player $j$ has \emph{at least} a probability $1/M$ to select
% a free arm, which implies $\overline{C^j(t+1)}$, and so implies the end of the sequence.
% %
% In other words, the average length of sequences of transitions $(2)$,
% $\E_{\mu}\left[ \tau_i - \theta_i \right]$,
% is bounded by the expected number of failed trial of a repeated Bernoulli experiment, with probability of success larger than $1/M$ (by the uniform choice of $A^j(t+1)$ in a set of size $M$ with at least one available arm).
% We recognize the mean of a geometric random variable, of parameter $\lambda \geq 1/M$, and so $\E_{\mu}\left[ \tau_i - \theta_i \right] = \frac{1}{\lambda} \leq \frac{1}{1/M} = M$.

% This finishes the proof as $\E_{\mu}[N_2^j(T)] = \sum_{t=1}^T \Pr(I_2^j(t)) = \bigO{\log T}$ and so
% $\sum_{t=1}^T \Pr(\widetilde{C^j}(t) \cap (A^j(t) = k)) = \bigO{\log T}$
% and finally
% $\E_{\mu}[\cC(T)] = \sum_{k=1}^K \E_{\mu}[\cC^k(T)] = \bigO{\log T}$ also.

% We can be more precise about the constants, all the previous arguments can be used successively:
% \begin{align}
%   \E_{\mu}[\cC(T)]
%   &\leq M \sum_{j=1}^M \left(\sum_{t=1}^T \Pr(I_2^j(t)) + \sum_{t=1}^T \Pr(I_3^j(t))\right)
%   = M \left(\sum_{j=1}^M \E_{\mu}[N_2^j(T)] + \E_{\mu}[N_3^j(T)]\right) \\
%   &\leq M^2 \left(\E_{\mu}[n(T)] \E_{\mu}[\theta_i - \tau_i] \right) + M^2 \E_{\mu}[N_3^1(T)] \notag  \\
%   &\leq M^2 (1+\E_{\mu}[N_3^1(T)] + \E_{\mu}[N_5^1(T)]) M + M^2 \E_{\mu}[N_3^1(T)] \notag \\
%   &\leq 2 M^3 \E_{\mu}[N_3^1(T)] + \smallO{\log T} + M^2 \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T} \notag  \\
%   &\leq \left(2 M^3 + M^2\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
% \end{align}

% And so we obtain the desired inequality, with explicit constants, that depend only on $\boldsymbol{\mu}$ and $M$.
% %
% \begin{equation}
%   \sum_{k=1}^K \E_{\mu}[\cC^k(T)] = \E_{\mu}[\cC(T)]
%   \leq M^2\left(2 M + 1\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
% \end{equation}

% \paragraph{Number of switches}\label{app:5:NumberSwitches}
% %
% Note that we controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
% which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
% Thus, the total number of arm switches is also proved to be logarithmic, if all players uses
% the \MCTopM-\klUCB{} algorithm.

% \paragraph{Strong uniform efficiency}\label{app:5:JustifyingDefinition5}
% %
% As soon as $R_T = \bigO{\log T}$ for all problem, \MCTopM{} is clearly proved
% to be uniformly efficient, as $\log T$ is $\smallO{T^{\alpha}}$ for any $\alpha\in(0,1)$.
% %
% And as justified after Definition~\ref{def:5:DecentralizedUniformEfficiency} (page~\pageref{def:5:DecentralizedUniformEfficiency}), uniform efficiency and invariance under permutations of the users implies strong uniform efficiency, and so \MCTopM{} satisfies Definition~\ref{def:5:DecentralizedUniformEfficiency}.
% %
% This is a sanity check: the lower-bound of
% Theorem~\ref{thm:5:BetterLowerBound} indeed applies to our algorithm \MCTopM,
% and finally this highlights that it is order-optimal for the regret, in the sense that it matches the lower-bound up-to a multiplicative constant,
% and optimal for the term \ref{eq:5:term1}.

% % -----------------------------------------------------------------
% % -----------------------------------------------------------------
% \subsection{Additional discussions on \Selfish}
% \label{app:5:SelfishFails}

% \TODOL{Maybe we can completely remove this, and just points to the paper \cite{Besson2018ALT}?}

% % OBSERVATION
% % Empirically it works fine in average of regrets on lots of simulations, but on worst case regret (or in a histogram) we can see it does not have a logarithmic average regret
% %
% % BAD
% % Explain on a small example ($K=2, M=3$ or $K=M=2$) that we found both manually, empirically and with a formal game tree exploration a small probability of failure, for both \Selfish-\UCB{} and \Selfish-\klUCB...
% %
% % a small probability of failure... $K=2, M=3$ or $K=M=2$. (idée: comparer \Selfish{} et \rhoRand{} en terme d'histogramme de la distribution de $R_T$)
% %
% % DONE estimer la proba ? ou la calculer avec le calcul formel ? ca c'est fait c'est bon
% %
% % a curiosity: when $K=M=2$ and $\mu_1 = \mu_2$, using \Selfish{} KL-UCB does orthogonalize !

% As said before, analyzing \Selfish{} is harder,
% but for instance one can prove that it yields constant
% collisions and regret for the trivial case of $\mu_1=\mu_2=1$ and $M=2$.
% %
% Empirically, when \Selfish{} is compared to the other algorithms,
% % with \UCB{} or \klUCB{} indices,
% % and the average regret is considered (for let say $1000$ repetitions),
% it is hard to find a case when \Selfish{} performs badly, as its (empirical average) regret always appeared logarithmic.
% %
% But an issue of only visualizing the empirical average regret for a certain number of repetitions is
% that if a certain ``bad'' run happens only with small probability, it is possible that it never happened
% in a simulation, or that it happened a few times but not enough to make the average regret look non-logarithmic\footnote{For instance, $0.999 \log(T) + 0.001 T$ looks more like $\log(T)$ than $T$ so an event yielding linear regret with ``small'' probability $10^{-3}$ cannot be observed from a plot showing the average regret.}.
% %
% This is why the distribution of regret, at the end of the simulations, $R_T$, is also displayed (in Appendix~\ref{app:5:moreplots}).
% %
% In a simple problem, with $M=2$ or $M=3$ players competing for $K=3$ arms,
% for instance with means $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$,
% the histogram in Figure~\ref{fig:5:selfish_fail1} shows that with a small probability, the regret $R_T$ of \Selfish-\klUCB{} is not small (and appears linear).
% %
% %
% Additionally, Figure~\ref{fig:5:selfish_fail2} shows that \Selfish-\klUCB{} also has bad performance against random uniform problems $\boldsymbol{\mu}\in[0,1]^K$,
% for $M=2$ or $3$ and $K=3$, in a lot of cases.
% In comparison, the others algorithms seem to have a logarithmic regret (for $M=2$ and all algorithms)
% or even a constant regret (for $M=3$ and \RandTopM{} and \MCTopM).

% The intuition behind these configurations when \Selfish{} performs poorly is the following,
% if all players use the same algorithm and the same indices.
% If two players $i$ and $j$ have exactly the same vectors $[\widetilde{S_k}^i(t)] = [\widetilde{S_k}^j(t)]$
% and $[T_k^i(t)] = [T_k^j(t)]$ at some time step $t \geq 1$,
% with different values for each $k$
% and so that the index vectors $\mathbf{g}^j(t) = \mathbf{g}^{i}(t)$ have different values for each arm $k$,
% then both players will take the same decisions at time $t$, and collide.
% Colliding does not change $[\widetilde{S_k}^j(t+1)]$ but increase one value in $[T_k^j(t+1)]$ by $1$.
% Then at the next step the same conditions on $\widetilde{S}^j$ and $N^j$ are preserved,
% and if the same condition on $\mathbf{g}^j(t+1)$ are also preserved,
% the two players will continue to collide.
% %
% We did not succeed in proving mathematically
% that the preservation of the first hypothesis on $\widetilde{S}^j$ and $N^j$
% implies the preservation of the hypothesis on index $\mathbf{g}^j$,
% but numerically it turns out to be always the case:
% and so two players colliding in such a setting will continue to do so infinitely:
% we denote such configurations as \emph{absorbing}.

% We wrote a script
% \footnote{~See online \texttt{\url{smpybandits.github.io/docs/complete_tree_exploration_for_MP_bandits.html}}}
% that explores formally all the possible runs,
% up-to a certain small time horizon,
% by exploring the complete \emph{game tree},
% of possible (random) rewards from the $K$ arms
% and (random) actions from the $M$ players, up-to a small depth of let say $T=8$.
% %
% Such game tree becomes quickly very large, but this was enough to confirm that with
% a certain small probability, function of $\mu_1,\dots,\mu_K$,
% two players can arrive in just a few steps in a ``bad'' absorbing configuration.
% %
% For instance, for only $K=2$ arms, the following game tree in Figure~\ref{fig:5:oneGameTree_SelfishKLUCB} illustrates the first $3$ steps that can lead to $2$ absorbing configurations.
% Using symbolic computations, the probability of reaching any of them was found to be
% $\geq \mu_1^2(1-\mu_2)^2/2 + \mu_2^2(1-\mu_1)^2/2$ for \Selfish-\UCB.
% That is far from being negligible, as it evaluates to $0.328$ for $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$,
% and a numerical simulation on $1000$ runs found $325$ cases of bad performance.
% %
% The same game tree exploration can be made for \Selfish-\klUCB,
% but so far we were not able to justify why it experiences fewer cases of bad performances even though our software found the same (lower bound on) failure probability.

% \begin{figure}[h!]
%     \centering
%     % \input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K2_M2_depth3__Selfish_UCB__absorbing__onlyfig.gv.tex}
%     \includegraphics[width=0.70\textwidth]{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing.pdf}
%     % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing__leafs.pdf}
%     % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing.pdf}
%     % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing__leafs.pdf}
%     \caption[Trajectories of $M=2$ players using \Selfish-\UCB{} for $K=2$ arms.]{For $K=2$ arms and $M=2$ players using \Selfish-\UCB, for depth$=3$: $2$ \textbf{\textcolor{red}{absorbing configurations}}. Each rectangle represents a configuration, as the matrix $[[\widetilde{S_k}^j(t) / T_k^j(t) ]_j]_k$. Absorbing configurations from depth $2$ are case of equality of the two vectors and the \Selfish{} indices $\widetilde{g_k}^j(t)$. Transitions are labeled with their probabilities.}
%     \label{fig:5:oneGameTree_SelfishKLUCB}
% \end{figure}

% From the structure of such game tree, we conjecture that the probability of reaching absorbing configurations (before a certain time $t$) is always lower-bounded
% by a polynomial function of $\mu_1,\dots,\mu_K$ and $1-\mu_1,\dots,1-\mu_K$,
% of degree at most $t$ in each variable.
% % with coefficients smaller than $1$.
% % and without simple factors (\ie, $\mu_k^d$).
% %
% As such, the lower bound on probability of failures should decrease when $K$ and $M$ increase, and this is coherent with the experiments for $K=9$ or $K=17$ (see Figures~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos} and \ref{fig:5:MP__K17_M6-12-17_T10000_N100__4_algos}),
% where \Selfish{} is shown to be uniformly more efficient than \rhoRand.
% %
% Of course, one cannot run an infinite number of simulations, and the smaller the probability of failure, the less likely it is to observe a failure
% in a finite number of runs.

% % IDEA
% \paragraph{Ideas to fix \Selfish{} ?}
% %
% It could be possible to change the \Selfish{} algorithm to add a way to escape such absorbing trajectories.
% For instance one could imagine that after seen seeing, \eg, $x=10$ collisions in a row,
% a certain random action could be taken by the players.
% These tricks can work empirically in some cases,
% but they are harder to analyze formally,
% and it is hard to tune the parameters (here $x$, but possibly more),
% and we do not find such tricks to be promising from a theoretical point-of-view.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\subsection{Additional figures}
\label{app:5:moreplots}

We include here plots missing from Section~\ref{sec:5:experiments},
and some additional numerical results.

% -----------------------------------------------------------------
\subsubsection{Illustration of the lower bound}
\label{app:5:illustrationLowerBound}

\TODOL{Maybe we can not include this part and just points to the paper \cite{Besson2018ALT}?}

We proved in Theorem~\ref{thm:5:BetterLowerBound} that the normalized regret, \ie, $R_T$ divided by
$\log T$, is asymptotically lower bounded by a constant $\mathrm{LB}(\boldsymbol{\mu}, M)$
depending on the problem $\boldsymbol{\mu}$ and the number of players $M$, for any $\rho$.
\begin{equation}
  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \rho)}{\log T} \geq \mathrm{LB}(\boldsymbol{\mu}, M).
\end{equation}
%
For an example problem with $K = 9$ arms, we display below on the $x$ axis is
the number of player, from $1$ player to $9$ players, and on the
$y$ axis is the value of this constant $\mathrm{LB}(\boldsymbol{\mu}, M)$, from the initial
theorem and from our theorem.
We chose a simple problem, with Bernoulli
distributed arms, with $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$.
% (as used in some articles).
%
Figure~\ref{fig:5:CompLowerBounds} clearly shows that our improved lower bound is indeed larger than the initial one by \cite{Zhao10},
and both become uninformative when $M=K$ (\ie, null).

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.85\textwidth]{Comparison_of_our_lowerbound_and_Anandkumar_2010__9_arms.pdf}
  \caption[Comparison of our lower bound against the one from \cite{Zhao10}]{Comparison of our lower bound against the one from \cite{Zhao10}, on a simple problem with $9$ Bernoulli arms, of means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$, as a function of the number of players $M$.}
  \label{fig:5:CompLowerBounds}
\end{figure}

%
% System regret and three terms
%
\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_2092905764868974160.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_1699475021767911583.pdf}
    % \end{subfigure}
    \caption[Multi-player regret with its three terms]{Regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
    \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

Figures~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}
show the regret $R_T(\boldsymbol{\mu}, M, \rho)$
on the same example problem $\boldsymbol{\mu}$, with $K = 9$ arms and respectively $M = 6$, or $9$ players, for \Selfish-\klUCB.
%
It is just a simple way to check that the two lower bounds on the regret indeed appear as valid lower bounds empirically,
and are moreover lower bounds on the count of selections (\ref{eq:5:term1}, displayed in \textcolor{blue}{cyan}).
%
The lower bounds (in black) are $C(\boldsymbol{\mu}, M) \log t$, the dashed line
for the lower bound from \cite{Zhao10}, and the continuous line is our lower bound.
%
These plot show the regret (in red),
% the three lower bounds
% (centralized, \citeauthor{Anandkumar11}'s and our lower bounds, in black),
and the three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3} in the decomposition of the regret.
As explained in Lemma~\ref{lem:5:DecompositionRegret}, term \ref{eq:5:term2} is not always non-negative.
For $M=9$ and \Selfish, \ref{eq:5:term3} is actually larger than the regret,
and term \ref{eq:5:term1} is zero, as well as the lower bounds.

% \subsubsection{Figures from Section~\ref{sec:5:experiments}}
% \label{app:5:plotsFromSec5}

% This last Appendix includes the figures used in Section~\ref{sec:5:experiments},
% with additional comments.