### 2019-05-27 4-Listofpublications/Listofpublications.tex
...

%!TEX root = ../PhD_thesis__Lilian_Besson
% ******************************* Thesis Appendix A ****************************
\chapter*{List of Publications}
\addcontentsline{toc}{chapter}{List of Publications}

% \TODOL{J'ai déjà inclus une liste de publication à la fin du Chapitre 1, nécessaire ici aussi ?}

% \section*{International Journals}
% \renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
% \begin{etaremune}%[label={[RI-\arabic*]}]
% 	\item \textbf{L. Besson}, E. Kaufmann ``XXX'',, in \emph{XXXX}, 2018 (Major Revision)
% \end{etaremune}

\section*{International Conferences}
\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
\begin{etaremune}
\item
  \href{https://hal.inria.fr/hal-XXX/document}{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation}, by \href{https://moychris.wordpress.com/}{C. Moy} \& \textbf{L. Besson},
  \href{https://sites.google.com/view/ISIoT2019/}{ISIoT Workshop} at \href{https://dcoss.org/}{DCOSS}, Santorini, Greece, May 2019.
  % \href{https://hal.inria.fr/hal-XXX}{HAL-XXX}.
  \cite{MoyBesson2019}
\item
  \href{https://perso.crans.org/besson/articles/BMBBM__IEEE_WCNC__2019.pdf}{Upper-Confidence
  Bound for Channel Selection in LPWA Networks with Retransmissions},
  by \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, \textbf{L. Besson}, J.
  Manco-Vasquez,
  \href{https://moychris.wordpress.com/}{C. Moy}.
  \href{https://sites.google.com/view/wcncworkshop-motion2019/}{MoTION Workshop} at
  \href{http://wcnc2019.ieee-wcnc.org/}{IEEE WCNC},
  Marrakech, Maroc, April 2019, \href{https://hal.inria.fr/hal-02049824}{HAL-02049824}.
  \cite{Bonnefoi2019WCNC}
\item
  \href{https://perso.crans.org/besson/articles/BBM__Demo_ICT_2018.pdf}{MALIN:
  « Multi-Arm bandit Learning for Iot Networks » with GRC: A TestBed
  Implementation and Demonstration that Learning Helps}, by \textbf{L.
  Besson}, \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi},
  \href{https://moychris.wordpress.com/}{C. Moy}. Demonstration
  presented at ICT (\emph{International Conference on Communication}),
  Saint-Malo, France, June 2018. Cf. \texttt{YouTu.be/HospLNQhcMk}, \href{https://hal.inria.fr/hal-02013866}{HAL-02013866}.
  \cite{Besson2018ICT}
\end{etaremune}

\section*{International Conferences}
\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
\begin{etaremune}
\item
  \href{https://perso.crans.org/besson/articles/BBM__IEEE_WCNC_2019.pdf}{GNU
  Radio Implementation of MALIN: « Multi-Armed bandits Learning for
  Internet-of-things Networks »}, by \textbf{L. Besson},
  \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi},
  \href{https://moychris.wordpress.com/}{C. Moy}.
  \href{http://wcnc2019.ieee-wcnc.org/}{IEEE WCNC},
  Marrakech, Maroc, April 2019, \href{https://hal.inria.fr/hal-02006825}{HAL-02006825}.
  \cite{Besson2019WCNC}
\item
  \href{https://hal.inria.fr/hal-01629733/document}{Multi-Player Bandits
  Revisited}, by \textbf{L. Besson} \&
  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
  Kaufmann}.
  \href{http://www.cs.cornell.edu/conferences/alt2018/accepted.html}{ALT}
  (\emph{Algorithmic Learning Theory}), Lanzarote, Spain, April 2018,
  \href{https://hal.inria.fr/hal-01629733}{HAL-01629733}.
  \cite{Besson2018ALT}
\item
  \href{https://hal.inria.fr/hal-01705292/document}{Aggregation of
  Multi-Armed Bandits learning algorithms for Opportunistic Spectrum
  Access}, by \textbf{L. Besson},
  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
  Kaufmann} \& \href{https://moychris.wordpress.com/}{C. Moy}.
  \href{http://wcnc2018.ieee-wcnc.org/}{IEEE WCNC}, Barcelone, Spain,
  avril 2018, \href{https://hal.inria.fr/hal-01705292}{HAL-01705292}.
  \cite{Besson2018WCNC}
\item
  \href{https://hal.inria.fr/hal-01575419/document}{Multi-Armed Bandit
  Learning in IoT Networks and non-stationary settings}, by
  \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, \textbf{L.
  Besson}, \href{https://moychris.wordpress.com/}{C. Moy},
  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
  Kaufmann} \& J. Palicot. \href{http://crowncom.org/2017/}{CrownCom}
  (\emph{Conference on Cognitive Radio Oriented Wireless Networks}),
  Lisbonne, Portugal, September 2017,
  \href{https://hal.inria.fr/hal-01575419}{HAL-01575419}, \textbf{Best Paper Award}.
  \cite{Bonnefoi17}
\end{etaremune}

\section*{National Conferences}
\renewcommand{\labelenumi}{[NC-\arabic{enumi}]}
\begin{etaremune}
\item
  \href{https://hal.inria.fr/hal-02006471/document}{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires}, by \textbf{L. Besson} \&
  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
  Kaufmann}.
  GRETSI, Lille, August 2019,
  \href{https://hal.inria.fr/hal-02006471}{HAL-02006471}.
  \cite{Besson2019Gretsi}
\end{etaremune}

\section*{Other National Communications}
\renewcommand{\labelenumi}{[ONC-\arabic{enumi}]}
\begin{etaremune}
  \item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, July 2017, \href{https://hal.inria.fr/hal-02013847}{HAL-02013847}.
  % \item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, March 2017.
  \item R. Bonnefoi, \textbf{L. Besson}, ``Multi-Armed Bandit Learning in IoT Networks'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, June 2018, \href{https://hal.inria.fr/hal-02013839}{HAL-02013839}.
\end{etaremune}

\section*{Other Publications}
\renewcommand{\labelenumi}{[OP-\arabic{enumi}]}
\begin{etaremune}
\item
  \href{https://hal.inria.fr/hal-02006471/document}{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits}, by \textbf{L. Besson} \&
  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
  Kaufmann}, February 2019,
  \href{https://hal.inria.fr/hal-02006471}{HAL-02006471}.
  \cite{Besson2019GLRT}
\item
  \href{https://hal.inria.fr/hal-01840022/document}{SMPyBandits: an
  Open-Source Research Framework for Single and Multi-Players Multi-Arms
  Bandits (MAB) Algorithms in Python}, by \textbf{L. Besson},
  active development since 2016,
  \href{https://hal.inria.fr/hal-01840022}{HAL-01840022}. Code on
  \href{https://GitHub.com/SMPyBandits/SMPyBandits}{GitHub.com/SMPyBandits}
  (about 40000 lines), documentation and tutorials on
  \href{https://SMPyBandits.rtfd.io}{SMPyBandits.rtfd.io}.
  \cite{SMPyBandits,SMPyBanditsJMLR}
\item
  \href{https://hal.inria.fr/hal-01736357/document}{What Doubling-Trick
  Can and Can't Do for Multi-Armed Bandits}, by \textbf{L. Besson} \&
  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
  Kaufmann}, September 2018,
  \href{https://hal.inria.fr/hal-01736357}{HAL-01736357}.
  \cite{Besson2018DoublingTricks}
\item
  \href{https://hal.inria.fr/hal-01847480/document}{A Note on the Ei
  Function and a Useful Sum-Inequality}, by \textbf{L. Besson}, February
  2018 \href{https://hal.inria.fr/hal-01847480}{HAL-01847480}.
\end{etaremune}

% \section*{Patents}
% \renewcommand{\labelenumi}{[PT-\arabic{enumi}]}
% \begin{etaremune}
% \item \textbf{L. Besson}, E. Kaufmann. ``XXX'', (pending)
% \end{etaremune}