### 2019-09-12 2-Chapters/5-Chapter/ALT_2018__MPBandits/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing__leafs__onlyfig.gv.tex
Removed raw inclusion of https://hal.inria.fr/hal-01629733v2

\begin{tikzpicture}[>=latex,line join=bevel,]
%%
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (302.0bp,11.5bp) node {Tree exploration for K=2 arms and M=2 players using Selfish KLUCB, for depth=3 : 68 leafs, 2 absorbing};
\end{scope}
  \node (0) at (247.1bp,128.0bp) [draw=green,ellipse] {[[0/0,0/0], [0/0,0/0]]};
  \node (19) at (134.1bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (68) at (361.1bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \draw [black,->] (0) ..controls (159.85bp,113.55bp) and (145.0bp,105.2bp)  .. (135.1bp,92.0bp) .. controls (130.31bp,85.627bp) and (128.83bp,77.397bp)  .. (19);
  \draw (237.1bp,84.5bp) node {$\mu_{2}^{2}(\mu_{1} - 1)^{2}/2$};
  \draw [black,->] (0) ..controls (315.21bp,109.06bp) and (328.69bp,101.89bp)  .. (339.1bp,92.0bp) .. controls (345.72bp,85.708bp) and (350.47bp,77.073bp)  .. (68);
  \draw (451.1bp,84.5bp) node {$\mu_{1}^{2}(\mu_{2} - 1)^{2}/2$};
%
\end{tikzpicture}