### 2019-09-12 2-Chapters/5-Chapter/ALT_2018__MPBandits/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing__onlyfig.gv.tex

\begin{tikzpicture}[>=latex,line join=bevel,]
%%
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (383.6bp,11.5bp) node {Tree exploration for K=2 arms and M=2 players using Selfish KLUCB, for depth=3 : 68 leafs, 2 absorbing};
\end{scope}
  \node (155) at (451.19bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \node (10) at (635.19bp,215.0bp) [draw,ellipse] {[[0/0,1/1], [0/1,0/0]]};
  \node (3) at (267.19bp,215.0bp) [draw,ellipse] {[[0/0,0/1], [1/1,0/0]]};
  \node (6) at (451.19bp,215.0bp) [draw,ellipse] {[[1/1,0/0], [0/0,0/1]]};
  \node (14) at (83.193bp,128.0bp) [draw=red,ellipse] {[[0/1,1/1], [0/1,1/1]]};
  \node (1) at (83.193bp,215.0bp) [draw,ellipse] {[[0/1,0/0], [0/0,1/1]]};
  \node (0) at (359.19bp,302.0bp) [draw=green,ellipse] {[[0/0,0/0], [0/0,0/0]]};
  \node (22) at (267.19bp,128.0bp) [draw=red,ellipse] {[[1/1,0/1], [1/1,0/1]]};
  \node (234) at (635.19bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (65) at (83.193bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (110) at (267.19bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \node (43) at (635.19bp,128.0bp) [draw=red,ellipse] {[[0/1,1/1], [0/1,1/1]]};
  \node (30) at (451.19bp,128.0bp) [draw=red,ellipse] {[[1/1,0/1], [1/1,0/1]]};
  \draw [red,->] (43) ..controls (635.19bp,98.163bp) and (635.19bp,82.548bp)  .. (234);
  \draw (645.69bp,84.5bp) node {$1$};
  \draw [black,->] (0) ..controls (225.73bp,289.45bp) and (154.26bp,280.04bp)  .. (128.19bp,266.0bp) .. controls (117.33bp,260.15bp) and (107.77bp,250.61bp)  .. (1);
  \draw (199.19bp,258.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (300.86bp,281.92bp) and (290.11bp,275.07bp)  .. (282.19bp,266.0bp) .. controls (276.58bp,259.57bp) and (273.06bp,251.1bp)  .. (3);
  \draw (353.19bp,258.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [red,->] (22) ..controls (267.19bp,98.163bp) and (267.19bp,82.548bp)  .. (110);
  \draw (277.69bp,84.5bp) node {$1$};
  \draw [black,->] (0) ..controls (485.53bp,296.17bp) and (538.83bp,287.27bp)  .. (582.19bp,266.0bp) .. controls (594.44bp,259.99bp) and (605.77bp,250.18bp)  .. (10);
  \draw (674.19bp,258.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (1) ..controls (83.193bp,185.16bp) and (83.193bp,169.55bp)  .. (14);
  \draw (149.19bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (10) ..controls (635.19bp,185.16bp) and (635.19bp,169.55bp)  .. (43);
  \draw (701.19bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (6) ..controls (451.19bp,185.16bp) and (451.19bp,169.55bp)  .. (30);
  \draw (517.19bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [red,->] (14) ..controls (83.193bp,98.163bp) and (83.193bp,82.548bp)  .. (65);
  \draw (93.693bp,84.5bp) node {$1$};
  \draw [red,->] (30) ..controls (451.19bp,98.163bp) and (451.19bp,82.548bp)  .. (155);
  \draw (461.69bp,84.5bp) node {$1$};
  \draw [black,->] (0) ..controls (406.04bp,280.6bp) and (416.21bp,273.97bp)  .. (424.19bp,266.0bp) .. controls (430.81bp,259.39bp) and (436.24bp,250.81bp)  .. (6);
  \draw (507.19bp,258.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (3) ..controls (267.19bp,185.16bp) and (267.19bp,169.55bp)  .. (22);
  \draw (333.19bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
%
\end{tikzpicture}