### 2019-09-12 2-Chapters/5-Chapter/ALT_2018__MPBandits/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__leafs__onlyfig.gv.tex

\begin{tikzpicture}[>=latex,line join=bevel,]
%%
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (13917.0bp,11.5bp) node {Tree exploration for K=2 arms and M=2 players using Selfish KLUCB, for depth=3 : 68 leafs, 2 absorbing};
\end{scope}
  \node (17) at (6572.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,1/2]]};
  \node (39) at (16256.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,0/1]]};
  \node (30) at (12781.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,2/2]]};
  \node (13) at (4517.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,1/1]]};
  \node (56) at (24250.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/2,0/1]]};
  \node (7) at (2786.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/2,1/1]]};
  \node (47) at (19388.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/2,0/1]]};
  \node (25) at (11345.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,1/1]]};
  \node (55) at (23905.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [0/2,0/1]]};
  \node (60) at (25419.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,0/1]]};
  \node (36) at (15327.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,1/2]]};
  \node (40) at (16604.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,2/2]]};
  \node (5) at (1344.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (65) at (26747.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,2/2]]};
  \node (43) at (18325.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,1/1]]};
  \node (2) at (351.19bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [0/2,0/1]]};
  \node (62) at (25894.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,2/2]]};
  \node (10) at (3643.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,1/2]]};
  \node (3) at (642.19bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,0/1]]};
  \node (67) at (27303.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,2/2]]};
  \node (19) at (7257.2bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (22) at (8945.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,2/2]]};
  \node (8) at (3027.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,1/1]]};
  \node (1) at (83.193bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,1/1]]};
  \node (20) at (7484.2bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,1/2]]};
  \node (59) at (25068.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,1/1]]};
  \node (35) at (14990.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,2/2]]};
  \node (0) at (14855.0bp,128.0bp) [draw=green,ellipse] {[[0/0,0/0], [0/0,0/0]]};
  \node (29) at (12437.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,1/2]]};
  \node (16) at (6230.2bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,0/2]]};
  \node (26) at (11613.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,0/1]]};
  \node (51) at (21910.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,2/2]]};
  \node (49) at (21443.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,0/1]]};
  \node (34) at (14721.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/1,0/2]]};
  \node (44) at (18608.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,0/1]]};
  \node (21) at (7832.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (64) at (26506.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,1/2]]};
  \node (9) at (3303.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,0/2]]};
  \node (24) at (10390.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,0/2]]};
  \node (52) at (22254.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/2,0/1]]};
  \node (18) at (6909.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,0/1]]};
  \node (31) at (13057.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (63) at (26162.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,0/1]]};
  \node (46) at (19048.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/2,0/1]]};
  \node (11) at (3983.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,1/1]]};
  \node (12) at (4251.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,0/1]]};
  \node (66) at (27015.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,0/2]]};
  \node (38) at (15916.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,1/1]]};
  \node (54) at (23637.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,2/2]]};
  \node (33) at (14438.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/1,1/2]]};
  \node (57) at (24516.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,0/2]]};
  \node (32) at (14170.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,1/2]]};
  \node (15) at (5954.2bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,1/2]]};
  \node (42) at (17988.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,1/1]]};
  \node (27) at (11901.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/1,1/2]]};
  \node (28) at (12100.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,0/1]]};
  \node (41) at (16895.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/2,0/1]]};
  \node (6) at (2444.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,1/1]]};
  \node (68) at (27654.0bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \node (23) at (9290.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (58) at (24799.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,1/2]]};
  \node (4) at (993.19bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,1/2]]};
  \node (53) at (22537.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (37) at (15675.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/1,1/2]]};
  \node (61) at (25695.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,2/2]]};
  \node (14) at (4861.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,0/2]]};
  \node (48) at (20343.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (50) at (21711.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,1/1]]};
  \node (45) at (18849.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,1/1]]};
  \draw [black,->] (0) ..controls (15310.0bp,123.9bp) and (16807.0bp,112.39bp)  .. (16851.0bp,92.0bp) .. controls (16863.0bp,86.743bp) and (16872.0bp,76.987bp)  .. (41);
  \draw (17406.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})(\mu_{1}^{2}\mu_{2}^{2} - 2\mu_{1}^{2}\mu_{2} + \mu_{1}^{2} - 2\mu_{1}\mu_{2}^{2} + 4\mu_{1}\mu_{2} - 2\mu_{1} + \mu_{2}^{2} - 2\mu_{2} + 2)/8$};
  \draw [black,->] (0) ..controls (13590.0bp,127.68bp) and (212.52bp,133.93bp)  .. (124.19bp,92.0bp) .. controls (113.32bp,86.84bp) and (104.37bp,77.236bp)  .. (1);
  \draw (246.69bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{3}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (14395.0bp,124.58bp) and (12861.0bp,115.02bp)  .. (12817.0bp,92.0bp) .. controls (12807.0bp,86.74bp) and (12799.0bp,77.252bp)  .. (30);
  \draw (12944.0bp,84.5bp) node {$\mu_{1}\mu_{2}^{3}(\mu_{1} - 1)^{2}/4$};
  \draw [black,->] (0) ..controls (13931.0bp,125.57bp) and (6996.8bp,114.16bp)  .. (6950.2bp,92.0bp) .. controls (6939.3bp,86.829bp) and (6930.4bp,77.223bp)  .. (18);
  \draw (7112.7bp,84.5bp) node {$-3\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/4$};
  \draw [black,->] (0) ..controls (15861.0bp,126.07bp) and (24152.0bp,117.69bp)  .. (24207.0bp,92.0bp) .. controls (24218.0bp,86.86bp) and (24228.0bp,77.259bp)  .. (56);
  \draw (24348.0bp,84.5bp) node {$-\mu_{1}(1-\mu_{1})(1-\mu_{2})^{2}/4$};
  \draw [black,->] (0) ..controls (14086.0bp,126.55bp) and (9395.4bp,122.62bp)  .. (9331.2bp,92.0bp) .. controls (9320.3bp,86.819bp) and (9311.4bp,77.211bp)  .. (23);
  \draw (9869.7bp,84.5bp) node {$\mu_{1}(1-\mu_{2})(3\mu_{1}^{2}\mu_{2}^{2} - 6\mu_{1}^{2}\mu_{2} + 3\mu_{1}^{2} - 6\mu_{1}\mu_{2}^{2} + 12\mu_{1}\mu_{2} - 6\mu_{1} + 3\mu_{2}^{2} - 6\mu_{2} + 4)/8$};
  \draw [black,->] (0) ..controls (14322.0bp,126.17bp) and (12202.0bp,121.22bp)  .. (12141.0bp,92.0bp) .. controls (12130.0bp,86.783bp) and (12121.0bp,77.169bp)  .. (28);
  \draw (12298.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (0) ..controls (13739.0bp,126.85bp) and (3412.8bp,124.59bp)  .. (3344.2bp,92.0bp) .. controls (3333.3bp,86.837bp) and (3324.4bp,77.232bp)  .. (9);
  \draw (3502.7bp,84.5bp) node {$\mu_{1}\mu_{2}(\mu_{1}  1)^{2}(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (13876.0bp,125.96bp) and (6047.6bp,116.91bp)  .. (5995.2bp,92.0bp) .. controls (5984.3bp,86.832bp) and (5975.4bp,77.226bp)  .. (15);
  \draw (6121.7bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}(\mu_{2} - 1)^{2}/4$};
  \draw [black,->] (0) ..controls (16003.0bp,126.94bp) and (26899.0bp,125.49bp)  .. (26972.0bp,92.0bp) .. controls (26983.0bp,86.865bp) and (26993.0bp,77.265bp)  .. (66);
  \draw (27124.0bp,84.5bp) node {$\mu_{1}^{2}(1-\mu_{1})(1-\mu_{2})^{3}/4$};
  \draw [black,->] (0) ..controls (15693.0bp,127.24bp) and (21323.0bp,127.58bp)  .. (21400.0bp,92.0bp) .. controls (21411.0bp,86.851bp) and (21421.0bp,77.249bp)  .. (49);
  \draw (21542.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15547.0bp,125.46bp) and (19293.0bp,116.1bp)  .. (19345.0bp,92.0bp) .. controls (19356.0bp,86.838bp) and (19366.0bp,77.233bp)  .. (47);
  \draw (19830.0bp,84.5bp) node {$(\mu_{1}^{2}\mu_{2}^{2} - 2\mu_{1}^{2}\mu_{2} + \mu_{1}^{2} - 2\mu_{1}\mu_{2}^{2} + 4\mu_{1}\mu_{2} - 2\mu_{1} + \mu_{2}^{2} - 2\mu_{2} + 2)/8$};
  \draw [black,->] (0) ..controls (13695.0bp,127.09bp) and (2559.0bp,127.07bp)  .. (2485.2bp,92.0bp) .. controls (2474.3bp,86.838bp) and (2465.4bp,77.233bp)  .. (6);
  \draw (2644.7bp,84.5bp) node {$-\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (13968.0bp,127.9bp) and (7611.3bp,133.0bp)  .. (7525.2bp,92.0bp) .. controls (7514.3bp,86.827bp) and (7505.4bp,77.221bp)  .. (20);
  \draw (7687.7bp,84.5bp) node {$-3\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/4$};
  \draw [black,->] (0) ..controls (14428.0bp,123.92bp) and (13125.0bp,113.08bp)  .. (13088.0bp,92.0bp) .. controls (13079.0bp,86.747bp) and (13072.0bp,77.521bp)  .. (31);
  \draw (13633.0bp,84.5bp) node {$\mu_{1}\mu_{2}(9\mu_{1}^{2}\mu_{2}^{2} - 18\mu_{1}^{2}\mu_{2} + 9\mu_{1}^{2} - 18\mu_{1}\mu_{2}^{2} + 36\mu_{1}\mu_{2} - 18\mu_{1} + 9\mu_{2}^{2} - 18\mu_{2} + 10)/8$};
  \draw [black,->] (0) ..controls (13615.0bp,127.55bp) and (768.1bp,132.31bp)  .. (683.19bp,92.0bp) .. controls (672.32bp,86.84bp) and (663.37bp,77.235bp)  .. (3);
  \draw (847.19bp,84.5bp) node {$3\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15978.0bp,126.79bp) and (26394.0bp,124.05bp)  .. (26463.0bp,92.0bp) .. controls (26474.0bp,86.864bp) and (26484.0bp,77.264bp)  .. (64);
  \draw (26591.0bp,84.5bp) node {$\mu_{1}\mu_{2}^{2}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (14167.0bp,125.48bp) and (10482.0bp,116.33bp)  .. (10431.0bp,92.0bp) .. controls (10420.0bp,86.81bp) and (10411.0bp,77.201bp)  .. (24);
  \draw (10897.0bp,84.5bp) node {$(\mu_{1}^{2}\mu_{2}^{2} - 2\mu_{1}^{2}\mu_{2} + \mu_{1}^{2} - 2\mu_{1}\mu_{2}^{2} + 4\mu_{1}\mu_{2} - 2\mu_{1} + \mu_{2}^{2} - 2\mu_{2} + 2)/8$};
  \draw [black,->] (0) ..controls (15425.0bp,126.88bp) and (17875.0bp,124.64bp)  .. (17945.0bp,92.0bp) .. controls (17956.0bp,86.818bp) and (17966.0bp,77.21bp)  .. (42);
  \draw (18121.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (13955.0bp,128.05bp) and (7387.1bp,134.31bp)  .. (7298.2bp,92.0bp) .. controls (7287.3bp,86.828bp) and (7278.4bp,77.222bp)  .. (19);
  \draw (7400.2bp,84.5bp) node {$\mu_{2}^{2}(\mu_{1} - 1)^{2}/2$};
  \draw [black,->] (0) ..controls (15710.0bp,127.42bp) and (21588.0bp,129.08bp)  .. (21668.0bp,92.0bp) .. controls (21679.0bp,86.853bp) and (21689.0bp,77.25bp)  .. (50);
  \draw (21775.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}/8$};
  \draw [black,->] (0) ..controls (13912.0bp,125.72bp) and (6661.8bp,115.13bp)  .. (6613.2bp,92.0bp) .. controls (6602.3bp,86.83bp) and (6593.4bp,77.224bp)  .. (17);
  \draw (6770.2bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15988.0bp,126.86bp) and (26633.0bp,124.73bp)  .. (26704.0bp,92.0bp) .. controls (26715.0bp,86.865bp) and (26725.0bp,77.264bp)  .. (65);
  \draw (26846.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{3}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (13599.0bp,127.62bp) and (478.87bp,133.15bp)  .. (392.19bp,92.0bp) .. controls (381.32bp,86.84bp) and (372.37bp,77.235bp)  .. (2);
  \draw (526.19bp,84.5bp) node {$\mu_{2}^{2}(\mu_{1}  1)^{3}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15743.0bp,127.79bp) and (22124.0bp,132.13bp)  .. (22211.0bp,92.0bp) .. controls (22222.0bp,86.855bp) and (22232.0bp,77.253bp)  .. (52);
  \draw (22360.0bp,84.5bp) node {$\mu_{1}\mu_{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (13990.0bp,127.67bp) and (7955.1bp,131.0bp)  .. (7873.2bp,92.0bp) .. controls (7862.3bp,86.826bp) and (7853.4bp,77.219bp)  .. (21);
  \draw (8418.2bp,84.5bp) node {$\mu_{1}\mu_{2}(9\mu_{1}^{2}\mu_{2}^{2} - 18\mu_{1}^{2}\mu_{2} + 9\mu_{1}^{2} - 18\mu_{1}\mu_{2}^{2} + 36\mu_{1}\mu_{2} - 18\mu_{1} + 9\mu_{2}^{2} - 18\mu_{2} + 10)/8$};
  \draw [black,->] (0) ..controls (15827.0bp,125.84bp) and (23542.0bp,115.96bp)  .. (23594.0bp,92.0bp) .. controls (23605.0bp,86.859bp) and (23615.0bp,77.257bp)  .. (54);
  \draw (23736.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (13630.0bp,127.46bp) and (1116.9bp,131.28bp)  .. (1034.2bp,92.0bp) .. controls (1023.3bp,86.839bp) and (1014.4bp,77.235bp)  .. (4);
  \draw (1198.2bp,84.5bp) node {$3\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (13893.0bp,125.85bp) and (6321.9bp,116.12bp)  .. (6271.2bp,92.0bp) .. controls (6260.3bp,86.831bp) and (6251.4bp,77.225bp)  .. (16);
  \draw (6430.7bp,84.5bp) node {$-\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (13772.0bp,126.64bp) and (4088.6bp,122.62bp)  .. (4024.2bp,92.0bp) .. controls (4013.3bp,86.836bp) and (4004.4bp,77.23bp)  .. (11);
  \draw (4146.7bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15057.0bp,120.35bp) and (15267.0bp,110.45bp)  .. (15297.0bp,92.0bp) .. controls (15306.0bp,86.539bp) and (15313.0bp,77.444bp)  .. (36);
  \draw (15475.0bp,84.5bp) node {$-3\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (15760.0bp,127.97bp) and (22404.0bp,133.72bp)  .. (22494.0bp,92.0bp) .. controls (22505.0bp,86.856bp) and (22515.0bp,77.254bp)  .. (53);
  \draw (23052.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})(3\mu_{1}^{2}\mu_{2}^{2} - 6\mu_{1}^{2}\mu_{2} + 3\mu_{1}^{2} - 6\mu_{1}\mu_{2}^{2} + 12\mu_{1}\mu_{2} - 6\mu_{1} + 3\mu_{2}^{2} - 6\mu_{2} + 4)/8$};
  \draw [black,->] (0) ..controls (13800.0bp,126.47bp) and (4619.4bp,121.07bp)  .. (4558.2bp,92.0bp) .. controls (4547.3bp,86.835bp) and (4538.4bp,77.229bp)  .. (13);
  \draw (4718.7bp,84.5bp) node {$3\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15961.0bp,126.69bp) and (26052.0bp,123.08bp)  .. (26119.0bp,92.0bp) .. controls (26130.0bp,86.864bp) and (26140.0bp,77.263bp)  .. (63);
  \draw (26299.0bp,84.5bp) node {$3\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (16015.0bp,127.01bp) and (27186.0bp,126.3bp)  .. (27260.0bp,92.0bp) .. controls (27271.0bp,86.865bp) and (27281.0bp,77.265bp)  .. (67);
  \draw (27443.0bp,84.5bp) node {$3\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (13645.0bp,127.38bp) and (1465.8bp,130.27bp)  .. (1385.2bp,92.0bp) .. controls (1374.3bp,86.839bp) and (1365.4bp,77.234bp)  .. (5);
  \draw (1923.7bp,84.5bp) node {$\mu_{1}(1-\mu_{2})(3\mu_{1}^{2}\mu_{2}^{2} - 6\mu_{1}^{2}\mu_{2} + 3\mu_{1}^{2} - 6\mu_{1}\mu_{2}^{2} + 12\mu_{1}\mu_{2} - 6\mu_{1} + 3\mu_{2}^{2} - 6\mu_{2} + 4)/8$};
  \draw [black,->] (0) ..controls (14663.0bp,120.94bp) and (14477.0bp,111.7bp)  .. (14453.0bp,92.0bp) .. controls (14446.0bp,86.326bp) and (14442.0bp,77.543bp)  .. (33);
  \draw (14583.0bp,84.5bp) node {$\mu_{1}\mu_{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (14303.0bp,126.64bp) and (12008.0bp,123.47bp)  .. (11942.0bp,92.0bp) .. controls (11931.0bp,86.788bp) and (11922.0bp,77.175bp)  .. (27);
  \draw (12030.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}/8$};
  \draw [black,->] (0) ..controls (13817.0bp,126.35bp) and (4961.2bp,120.07bp)  .. (4902.2bp,92.0bp) .. controls (4891.3bp,86.834bp) and (4882.4bp,77.229bp)  .. (14);
  \draw (5437.2bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})(\mu_{1}^{2}\mu_{2}^{2} - 2\mu_{1}^{2}\mu_{2} + \mu_{1}^{2} - 2\mu_{1}\mu_{2}^{2} + 4\mu_{1}\mu_{2} - 2\mu_{1} + \mu_{2}^{2} - 2\mu_{2} + 2)/8$};
  \draw [black,->] (0) ..controls (15504.0bp,124.85bp) and (18761.0bp,113.1bp)  .. (18806.0bp,92.0bp) .. controls (18817.0bp,86.832bp) and (18827.0bp,77.226bp)  .. (45);
  \draw (18913.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{3}/8$};
  \draw [black,->] (0) ..controls (15225.0bp,125.7bp) and (16158.0bp,120.08bp)  .. (16215.0bp,92.0bp) .. controls (16226.0bp,86.714bp) and (16235.0bp,77.089bp)  .. (39);
  \draw (16395.0bp,84.5bp) node {$-3\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (16034.0bp,127.11bp) and (27535.0bp,127.3bp)  .. (27611.0bp,92.0bp) .. controls (27622.0bp,86.866bp) and (27632.0bp,77.266bp)  .. (68);
  \draw (27732.0bp,84.5bp) node {$\mu_{1}^{2}(\mu_{2} - 1)^{2}/2$};
  \draw [black,->] (0) ..controls (15947.0bp,126.61bp) and (25785.0bp,122.32bp)  .. (25851.0bp,92.0bp) .. controls (25862.0bp,86.863bp) and (25872.0bp,77.263bp)  .. (62);
  \draw (25993.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{3}(1-\mu_{1})/8$};
  \draw [black,->] (0) ..controls (15924.0bp,126.46bp) and (25313.0bp,120.98bp)  .. (25376.0bp,92.0bp) .. controls (25387.0bp,86.863bp) and (25397.0bp,77.262bp)  .. (60);
  \draw (25522.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}(\mu_{2} - 1)^{2}/4$};
  \draw [black,->] (0) ..controls (15877.0bp,126.16bp) and (24416.0bp,118.43bp)  .. (24473.0bp,92.0bp) .. controls (24484.0bp,86.861bp) and (24494.0bp,77.26bp)  .. (57);
  \draw (24622.0bp,84.5bp) node {$\mu_{1}\mu_{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (14276.0bp,127.28bp) and (11727.0bp,126.73bp)  .. (11654.0bp,92.0bp) .. controls (11643.0bp,86.794bp) and (11634.0bp,77.182bp)  .. (26);
  \draw (11787.0bp,84.5bp) node {$\mu_{1}^{2}(1-\mu_{1})(1-\mu_{2})^{3}/4$};
  \draw [black,->] (0) ..controls (14252.0bp,127.82bp) and (11465.0bp,129.76bp)  .. (11386.0bp,92.0bp) .. controls (11375.0bp,86.798bp) and (11366.0bp,77.187bp)  .. (25);
  \draw (11509.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{3}(1-\mu_{1})/8$};
  \draw [black,->] (0) ..controls (15271.0bp,127.62bp) and (16487.0bp,126.88bp)  .. (16561.0bp,92.0bp) .. controls (16572.0bp,86.767bp) and (16582.0bp,77.151bp)  .. (40);
  \draw (16713.0bp,84.5bp) node {$\mu_{2}^{2}(\mu_{1}  1)^{3}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15905.0bp,126.35bp) and (24964.0bp,119.99bp)  .. (25025.0bp,92.0bp) .. controls (25036.0bp,86.862bp) and (25046.0bp,77.261bp)  .. (59);
  \draw (25208.0bp,84.5bp) node {$3\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15128.0bp,127.97bp) and (15581.0bp,125.56bp)  .. (15641.0bp,92.0bp) .. controls (15651.0bp,86.628bp) and (15658.0bp,77.252bp)  .. (37);
  \draw (15765.0bp,84.5bp) node {$\mu_{1}\mu_{2}^{2}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (15843.0bp,125.94bp) and (23809.0bp,116.71bp)  .. (23862.0bp,92.0bp) .. controls (23873.0bp,86.86bp) and (23883.0bp,77.258bp)  .. (55);
  \draw (24042.0bp,84.5bp) node {$\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (14357.0bp,125.42bp) and (12529.0bp,118.08bp)  .. (12476.0bp,92.0bp) .. controls (12466.0bp,86.799bp) and (12457.0bp,77.321bp)  .. (29);
  \draw (12637.0bp,84.5bp) node {$3\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15520.0bp,125.08bp) and (18957.0bp,114.21bp)  .. (19005.0bp,92.0bp) .. controls (19016.0bp,86.834bp) and (19026.0bp,77.229bp)  .. (46);
  \draw (19183.0bp,84.5bp) node {$\mu_{1}\mu_{2}(\mu_{1}  1)^{2}(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (13725.0bp,126.93bp) and (3138.5bp,125.39bp)  .. (3068.2bp,92.0bp) .. controls (3057.3bp,86.837bp) and (3048.4bp,77.232bp)  .. (8);
  \draw (3194.7bp,84.5bp) node {$\mu_{1}\mu_{2}^{3}(\mu_{1} - 1)^{2}/4$};
  \draw [black,->] (0) ..controls (15456.0bp,127.57bp) and (18204.0bp,128.37bp)  .. (18282.0bp,92.0bp) .. controls (18293.0bp,86.824bp) and (18303.0bp,77.217bp)  .. (43);
  \draw (18431.0bp,84.5bp) node {$\mu_{1}\mu_{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (13754.0bp,126.74bp) and (3750.7bp,123.6bp)  .. (3684.2bp,92.0bp) .. controls (3673.3bp,86.836bp) and (3664.4bp,77.231bp)  .. (10);
  \draw (3842.7bp,84.5bp) node {$\mu_{1}\mu_{2}(\mu_{1}  1)^{2}(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (14760.0bp,111.37bp) and (14731.0bp,103.07bp)  .. (14722.0bp,92.0bp) .. controls (14717.0bp,85.643bp) and (14716.0bp,77.29bp)  .. (34);
  \draw (14845.0bp,84.5bp) node {$\mu_{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15937.0bp,126.55bp) and (25588.0bp,121.76bp)  .. (25652.0bp,92.0bp) .. controls (25663.0bp,86.863bp) and (25673.0bp,77.262bp)  .. (61);
  \draw (25759.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{3}/8$};
  \draw [black,->] (0) ..controls (14605.0bp,127.17bp) and (14241.0bp,123.3bp)  .. (14194.0bp,92.0bp) .. controls (14186.0bp,86.644bp) and (14181.0bp,77.814bp)  .. (32);
  \draw (14317.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15618.0bp,126.37bp) and (20237.0bp,121.43bp)  .. (20300.0bp,92.0bp) .. controls (20311.0bp,86.845bp) and (20321.0bp,77.242bp)  .. (48);
  \draw (20858.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})(3\mu_{1}^{2}\mu_{2}^{2} - 6\mu_{1}^{2}\mu_{2} + 3\mu_{1}^{2} - 6\mu_{1}\mu_{2}^{2} + 12\mu_{1}\mu_{2} - 6\mu_{1} + 3\mu_{2}^{2} - 6\mu_{2} + 4)/8$};
  \draw [black,->] (0) ..controls (13784.0bp,126.56bp) and (4355.0bp,121.84bp)  .. (4292.2bp,92.0bp) .. controls (4281.3bp,86.835bp) and (4272.4bp,77.23bp)  .. (12);
  \draw (4413.7bp,84.5bp) node {$-\mu_{1}(1-\mu_{1})(1-\mu_{2})^{2}/4$};
  \draw [black,->] (0) ..controls (14938.0bp,111.65bp) and (14955.0bp,103.82bp)  .. (14968.0bp,92.0bp) .. controls (14975.0bp,85.829bp) and (14980.0bp,77.081bp)  .. (35);
  \draw (15136.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (0) ..controls (15722.0bp,127.56bp) and (21785.0bp,130.2bp)  .. (21867.0bp,92.0bp) .. controls (21878.0bp,86.853bp) and (21888.0bp,77.251bp)  .. (51);
  \draw (22047.0bp,84.5bp) node {$3\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15480.0bp,128.1bp) and (18480.0bp,131.51bp)  .. (18565.0bp,92.0bp) .. controls (18576.0bp,86.829bp) and (18586.0bp,77.222bp)  .. (44);
  \draw (18693.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15174.0bp,123.27bp) and (15837.0bp,113.45bp)  .. (15878.0bp,92.0bp) .. controls (15888.0bp,86.683bp) and (15897.0bp,77.186bp)  .. (38);
  \draw (16053.0bp,84.5bp) node {$\mu_{1}\mu_{2}(\mu_{1}  1)^{2}(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (15891.0bp,126.26bp) and (24697.0bp,119.23bp)  .. (24756.0bp,92.0bp) .. controls (24767.0bp,86.862bp) and (24777.0bp,77.26bp)  .. (58);
  \draw (24898.0bp,84.5bp) node {$\mu_{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (14063.0bp,126.84bp) and (9054.6bp,124.6bp)  .. (8986.2bp,92.0bp) .. controls (8975.3bp,86.821bp) and (8966.4bp,77.213bp)  .. (22);
  \draw (9147.2bp,84.5bp) node {$\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (13713.0bp,126.99bp) and (2898.9bp,126.08bp)  .. (2827.2bp,92.0bp) .. controls (2816.3bp,86.837bp) and (2807.4bp,77.232bp)  .. (7);
  \draw (2936.2bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}(1-\mu_{2})/4$};
%
\end{tikzpicture}